---
title: "Week 3, Day 3 : R and SQL"
output: html_notebook
---

```{r}
library(RPostgres)
library(tidyverse)
```


# Connecting R to an SQL Database

## Storing Access Credentials Securely

Two strategies when working in Git -

- Don't store the credentials anywhere near the Git repository
- Use a .gitignore file

### .gitignore

1) modify / create .gitignore to ignore the file
2) commit the .gitignore
3) add file with credentials to project
4) source file into script/notebook

```{r}
source("secrets.R")
```

## Connecting

```{r}
omni_connect <- dbConnect(
  drv = Postgres(),
  dbname = database,
  host = host,
  port = port,
  user = username,
  password = password
)

rm(username, password)
```

# Things we can do with our connection

## Tables

```{r}
# see what tables are in the database
dbListTables(conn = omni_connect)
```

```{r}
# see what columns are in a table
dbListFields(conn = omni_connect, name = "employees")
```

```{r}
# lets us query the database
employees <- dbGetQuery(conn = omni_connect, statement = "SELECT * FROM employees LIMIT 10")
```

```{sql, connection = omni_connect, output.var = "highest_earners"}
/*
Querying in SQL code chunks
*/

SELECT *
FROM employees
ORDER BY salary DESC
LIMIT 100;
```

```{r}
highest_earners %>% 
  filter(department == "Accounting")
```

