---
title: "Strings"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Basics

Strings - R understands both single and double quotes for strings.

BE VERY CAREFUL if you decide to use single quotes. It'll fuck up in a lot of strings

'this is Dave's dog', for instance, will fall over. Stick to double quotes in most instances!

```{r}
employees <- tibble(
  first_name = c("Sally", "steven", "Mary", "John"),
  last_name = c("Smith", "hasTings", "Milne", "Davidson"),
  postcode = c("EH10 6GH", "EH3 7HL", "G1 7FR", NA), 
  location = c(" ", "Edinburgh", "", "Stirling")
)

employees
```

Strings can contain any sort of character. Different numbers of spaces are all unique strings, and one space is not the same as an empty string!

```{r}
employees %>% 
  group_by(location) %>% 
  summarise(count = n())
```

```{r}
mary_location <- employees %>% 
  filter(first_name == "Mary") %>% 
  select(location) %>% 
  pull()

sally_location <- employees %>% 
  filter(first_name == "Sally") %>% 
  select(location) %>% 
  pull()

mary_location == sally_location
```

Time for `stringr`

concatenate strings with two functions:
`str_c`
`str_join`

```{r}
str_c("This","is","some","text","that","I","wanttojoin")
```
```{r}
employees %>% 
  mutate(full_name = str_c(first_name, last_name, sep=' '))
```

```{r}
employees %>% 
  mutate(full_name = str_c(first_name, location, sep=' '))
```

# Casings

```{r}
pangram <- " Todays class is about basics of string"

pangram %>% 
  str_to_sentence()
```
```{r}
employees %>% 
  select(first_name, last_name) %>% 
  mutate(first_name_fixed = str_to_title(first_name)) %>% 
  mutate(last_name_fixed = str_to_title(last_name))
```

