---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

# Extra dplyr

 - `groupby()` with `mutate()`
 - using `across` to perform operations across multiple columns
 
 When group_by -> mutating the grouping level remains. In summarise the default
 is to remove the last layer of grouping

```{r}
chinesemeal %>% 
  group_by(Year) %>% 
  mutate(total_calories = sum(CaloriesPerDay)) %>% 
  ungroup() %>% # remove any persistent groups
  mutate(proportion_total_cals = ((CaloriesPerDay / total_calories))*100) %>% 
  arrange(Year)
```

# window functions

`lag()` and `lead()`

`lag()` accessess a value in a previous row.
Use it to get differences over time.

For Food Type = Carbohydrates, add a column showing the difference in CaloriesPerDay 
compared with the previous year. Note! You can give lag a value if you want to look more than one back

```{r}
chinesemeal %>% 
  filter(FoodType == "Carbohydrates") %>% 
  mutate(prev_cals = lag(CaloriesPerDay, 1)) %>% 
  mutate(calorie_diff = CaloriesPerDay - prev_cals)
```

# Grouping on multiple columns

Get a count of how many people in each allegiance are of each gender.
Note this example replicates `count()`

```{r}
all_deaths %>% 
  group_by(allegiances, gender) %>% 
  summarise(n_characters = n())
```

# Operations across multiple columns

Change column types of the book columns to be logical rather than double.
When running a function, `across()` is looking for the NAME of a function, not a function! no parenthese!

```{r}
all_deaths %>% 
  mutate(
    across(.cols = book1_GoT:book5_DwD,
           .fns = as.logical)
  )
```

Get the mean value of each numerical column
```{r}
all_deaths %>% 
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = mean, #can apply additional arguments to the function  
      na.rm = TRUE
    )
  )
```

