---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(broom)
library(pROC)
```

```{r}
mortgage_data <- read_csv("data/mortgage_applications.csv") %>% 
  janitor::clean_names()
```

```{r message=FALSE}
GGally::ggpairs(mortgage_data)
```

```{r}
mortgage_data %>% 
  ggplot(aes(x = tu_score, y = accepted)) +
  geom_col()

mortgage_data %>% 
  ggplot(aes(x = tu_score, y = accepted)) +
  geom_boxplot() +
  geom_point(position = position_jitter(height = .1))

mortgage_data %>% 
  ggplot() +
  geom_density(aes(x = tu_score))
```

- probability
  - defined as success divided by all possibilities
  - on a d6: 1/6
  - 0-1 measure of chance of event
  
- odds
  - defined as the probability of success divided by the probability of failure
  - on a d6: (1/6) / (5/6) = 1/5 or 5 to 1
  - the ratio of successes to fails
  
- odds ratio
  - Ratio of odds
  - odds of one event / odds of another event

- ln / log
  - natural log
  - A multiplicative scale

- log odds
  - literally the log of the odds
  - on a d6: log(1/5) or log(0.2)
  - What the logistic model returns

- link function
  - In "logit space" the y axis is in *log odds* units.
    We need a *link function* to map the logits to our units - which are probabilities. 

- logit 
  -   *log odds*
  
- odds factor

### Logistic Regression

`glm()` Generalised linear model

Know how to use the `glm()` function i nR
Carry out logistic regression in R

```{r}
mortgage_data_logreg_model <- glm(accepted ~ tu_score, mortgage_data,
                                  family = binomial)
```

```{r}
summary(mortgage_data_logreg_model)
```

Coefficients are in __logits__

$$
ln(odds_{success}(x)) = \beta_0 + \beta_1 x
$$

```{r}
(-4.575035 + 0.008475 * 600) %>% 
  exp()
```

```{r}
predict_log <- tibble(tu_score = seq(0, 710, 1)) %>% 
  add_predictions(mortgage_data_logreg_model, type = 'response')
```

```{r}
ggplot(mortgage_data) +
  geom_jitter(aes(x = tu_score, y = as.integer(accepted)),
              shape = 1,
              height = 0.05) +
  geom_line(aes(x = tu_score, y = pred), data = predict_log, col = "red", linewidth = 2) +
  labs(y = "Probability")
```

```{r}
odds_at_594 <- predict_log %>% 
  filter(tu_score == 594) %>% 
  mutate(odds_factor = pred / (1 - pred)) %>% 
  select(odds_factor)
```

odds of being accepted with a TU score of 594 = 1.58
How would those odds change if you score was 50 higher: 644

```{r}
b_tu_score <- tidy(mortgage_data_logreg_model) %>% 
  filter(term == "tu_score") %>% 
  select(estimate)

odds_factor <- exp(b_tu_score * 50)

odds_at_594

odds_at_644 <- odds_factor * odds_at_594

odds_at_644

predict_log %>% 
  filter(tu_score == 594 | tu_score == 644) %>% 
  mutate(odds = pred / (1 - pred)) 
```

```{r}
2.417739	- 1.582625
```


### Multiple predictors

$$
\alpha + (\beta_{score} \times {score}) + (\beta_{job} \times {jobstatus})
$$
```{r}
mortgage_data_multi_logreg_model <- glm(accepted ~ tu_score + employed,
                                        mortgage_data , family = binomial)

summary(mortgage_data_multi_logreg_model)

janitor::clean_names(tidy(mortgage_data_multi_logreg_model))
```

log odds of being accepted for a mortgage when employed are 1.48 higher
1.48 *logits* higher
to convert to odds we `exp()` it

```{r}
exp(1.484537909)
```
to convert from odds to probability:

1 / (1 - e^-lm)

```{r}
1 / (1 + exp(-(-4.686827492	+ 0.006723941 * 500))) # 500 score, no job
1 / (1 + exp(-(-4.686827492	+ 0.006723941 * 500 + 1.484537909))) # 500 score, job
```

```{r}
# odds of acceptance when employed

b_employed_true <- tidy(mortgage_data_multi_logreg_model) %>% 
  filter(term == "employedTRUE") %>% 
  select(estimate)

exp(b_employed_true)
```

Be able to check for statiscal significane of fitted coefficients via their p-values.

Be able to calculate a _confusion matrix_ to calculate accuracy, and true and false positive rates.

```{r}
mortgage_3pred_model <- glm(accepted ~ tu_score + employed + age, data = mortgage_data, family = binomial(link = 'logit'))

mortgage_data_with_3pred <- mortgage_data %>%
  add_predictions(mortgage_3pred_model, type = "response")
```

```{r}
threshold <- 0.6

mortgage_data_with_3pred <- mortgage_data_with_3pred %>% 
  mutate(pred_threshold_0.6 = pred >= threshold)


```

### Confusion Matrix

```{r}
conf_mat <- mortgage_data_with_3pred %>% 
  tabyl(accepted, pred_threshold_0.6) %>% 
  adorn_title()
```

## Accuracy

```{r}
conf_mat

(179 + 679) / 1000
```
Our model was 85.5% accurate

For acuracy to be at its best, we need a "balanced" data set

(nTP + nTN) / n

```{r}

```

