---
title: "dplyr verbs"
output: html_notebook
---

Goal today is to figure out what all the dplyr verbs do and how to use them

On the block today --

select()
filter()
arrange()
mutate()
group_by()
summarise()

First up, load libraries. CodeClanData is where all the class data is kept.

```{r}
library(tidyverse)
library(CodeClanData)
library(skimr)
```

Brief R lesson -- two ways to call functions from libraries. Either load the entitre library with the library() call and then use the function, or via calling the function explicitly via libraryname::function_name eg praise::praise()

CodeClanData has a whole bunch of datasets inside it, so we're going to use those.

```{r}
all_deaths
```

This is a very daft dataset. Game of Thrones!

How big is this dataset?
```{r}
nrow(all_deaths)
```

will let us know how many rows.

```{r}
ncol(all_deaths)
```

for columns.

Want the dimensions?

```{r}
dim(all_deaths)
```

will do that.

```{r}
names(all_deaths)
```

gives us the names of all the columns (and give us column numbers!)

```{r}
head(all_deaths)
```

will give us the top of the data. You can customise how many rows you see --

```{r}
head(all_deaths, 10)
```

you can use tail() to see the bottom instead

```{r}
tail(all_deaths)
```

Cool summary function -

```{r}
glimpse(all_deaths)
```

Want a little sanity? Have a look at the dataset as an actual table! Probably don't want this in code, but it might be useful in console!

```{r}
View(all_deaths)
```

Useful function in skimr:

```{r}
skim(all_deaths)
```
Gives you a pretty simple summation of the data.

Time for dplyr! select() will let you pull just the columns you want

```{r}
select(all_deaths, name, allegiances, gender, nobility, year_of_death)
```

You can also do the reverse and just nominate columns to not view.

```{r}
select(all_deaths, -name)
```

Reminder that without assigning these to an object, you're not actually doing anything besides view it.

```{r}
all_deaths_no_names <- select(all_deaths, -name)
```

will throw it into an object. If you want it to be a little more interactive, wrap the whole line in parentheses and RStudio will pop a view of the new object for you

```{r}
(all_deaths_no_names <- select(all_deaths, -name))
```

Cool shit you can do -- 

select() has a whole bunch of helper functions. Check the man page with ?

```{r}
?select
```

What if you wanted just the columns related to a book?

```{r}
select(all_deaths, contains("Book"))
```

how about related to a death?

```{r}
select(all_deaths, contains("Death"))
```

You can combine things to get useful stuff like this --

```{r}
select(all_deaths, name, contains("Death"))
```

Now for filter(). Same as select(), but for rows! Let's you return all rows that meet a condition.

Makes big use of the logical operators.

Useful one you might not know --

%in% checks for equality against *more than one* value.

Time for some examples -

```{r}
filter(all_deaths, allegiances == "Lannister")
```

```{r}
filter(all_deaths, allegiances != "Lannister")
```

Shows everyone who is or isn't a Lannister

```{r}
filter(all_deaths, allegiances %in% c("House Lannister", "Lannister"))
```

Note that this pulls more results than the initial filter! The filter is doing exact matches, so all the ones with the "House" prefix *were not included*

```{r}
filter(all_deaths, year_of_death >= 299)
```

Sidebar into a handy bonus function --

```{r}
distinct(all_deaths, allegiances)
```

will show you all the distinct values from a column. So if you're trying to glance at all the House names or something, this is a handy way to give it a glance.

Might be helpful to manipulate it a bit, so

```{r}
distinct_alleg <- distinct(all_deaths, allegiances)

arrange(distinct_alleg, allegiances)
```

Will put them all in distinct_alleg and then sort them alphabetically

```{r}
filter(all_deaths, allegiances == "Lannister" 
       & nobility == 1)
```

```{r}
filter(all_deaths, book_intro_chapter >= 5 &
         book_intro_chapter <= 10)
```

Quick note -- AND takes precedence over OR

```{r}
filter(all_deaths, year_of_death <= 299)
```


```{r}
filter(all_deaths, name %in% c("Jon Snow", "Samwell Tarly", "Daenerys Targaryen"))
```

```{r}
got_characters <- filter(all_deaths, allegiances != "Lannister" & gender == 0)
select(got_characters, name)
```
OR NEST THIS

```{r}
select(filter(all_deaths, allegiances != "Lannister" & gender == 0), name)
```

```{r}
all_deaths %>% 
  filter(allegiances != "Lannister" & gender == 0) %>%
  select(name)
```

PIPES OWN. CTRL+SHIFT+M %>% %>% %>% %>% %>% %>% %>% 

Time for arrange(). By default this is in ascending order

```{r}
all_deaths %>% 
  arrange(gender)
```

Wanna do it descending?

```{r}
all_deaths %>% 
  arrange(desc(gender))
```

You can sort my multiple columns. Sorts in priority from left to right.

```{r}
all_deaths %>% 
  arrange(book_of_death, death_chapter)
```

```{r}
all_deaths %>% 
  arrange(allegiances)
```

```{r}
all_deaths %>% 
  arrange(allegiances, book_intro_chapter)
```

```{r}
all_deaths %>% 
  arrange(desc(year_of_death))
```

mutate() up next! Use this to create or modify columns

```{r}
all_deaths %>% 
  mutate(years_survived = year_of_death - 298, .after = year_of_death) %>% 
  arrange(desc(years_survived))
```

this will make a new column years_survived which is a subtraction of years_of_death and 298. .after dictates where the column goes. .before also works!

Mutate can be used with conversion functions. Stuff like

as.integer()
as.double()
as.numeric()
as.character()
as.factor()
as.logical()

```{r}
all_deaths %>% 
  mutate(book_of_death = as.character(book_of_death))
```

BE CAREFUL. Converting in ways that don't make sense will blow up your stuff.

```{r}
all_deaths %>% 
  mutate(name = as.numeric(name))
```

Telling it to convert characters to numerics will N/A out data!

```{r}
all_deaths %>% 
  mutate(year_of_death = mean(year_of_death, na.rm = TRUE))
```

the na.rm in the above is a flag to tell the calculation to exclue N/As.

Quick task --

1 Add a new column which is a logical vector, which has TRUE if book_intro_chapter is empty, and FALSE if it isn’t. Hint: you’ll need to use the is.na function.
2 Change the nobility column to a logical vector.
3 Find the number of male characters by taking the sum of the gender column (as gender = 1 is male).

```{r}
all_deaths %>% 
  mutate(no_intro = as.logical(is.na(book_intro_chapter)))
```

```{r}
all_deaths %>% 
  mutate(as.logical(nobility))
```

```{r}
all_deaths %>% 
  mutate(men = sum(gender))
```

last one not very useful!

NEXT! summarise()

This is a better way to do question 3 from above. Few other options though!

```{r}
all_deaths %>% 
  summarise(n_males = sum(gender))
```

AND groupby()

```{r}
all_deaths %>% 
  group_by(allegiances) %>% 
  summarise(character_count = n())
```

```{r}
all_deaths %>% 
  group_by(allegiances) %>% 
  summarise(character_count = n())
```

n() function simply reports the size of a group, so it's taking those groups, counting their members and then summarising the result.

```{r}
all_deaths %>% 
  group_by(nobility, gender) %>% 
  summarise(character_count = n())
```

New task!

If a character doesn’t have a missing value for book_of_death, this means that they have died at some point. Use this to find the allegiances with the highest number of people dead.
Hint You will need to:
Filter the data so we just have characters who have died.
Use group by and summarise to find the number of characters who belonged to each allegiance.
Arrange so the highest values come first. 
Try to use a single pipe sequence for this solution.

```{r}
all_deaths %>% 
  filter(!is.na(book_of_death)) %>% #
  group_by(allegiances) %>% 
  summarise(character_count = n()) %>% 
  arrange(desc(character_count))
```

```{r}
all_deaths %>% 
  filter(is.na(book_of_death))
```

NEXT! pull()

```{r}
average_year_of_death <- all_deaths %>% 
  summarise(avg_value = mean(year_of_death, na.rm = TRUE)) %>% 
  pull()
all_deaths %>% 
  mutate(death_later_than_average = year_of_death > average_year_of_death)
```

