---
title: "dplyr verbs"
output: html_notebook
---

Goal today is to figure out what all the dplyr verbs do and how to use them

On the block today --

select()
filter()
arrange()
mutate()
group_by()
summarise()

First up, load libraries. CodeClanData is where all the class data is kept.

```{r}
library(tidyverse)
library(CodeClanData)
library(skimr)
```

Brief R lesson -- two ways to call functions from libraries. Either load the entitre library with the library() call and then use the function, or via calling the function explicitly via libraryname::function_name eg praise::praise()

CodeClanData has a whole bunch of datasets inside it, so we're going to use those.

```{r}
all_deaths
```

This is a very daft dataset. Game of Thrones!

How big is this dataset?
```{r}
nrow(all_deaths)
```

will let us know how many rows.

```{r}
ncol(all_deaths)
```

for columns.

Want the dimensions?

```{r}
dim(all_deaths)
```

will do that.

```{r}
names(all_deaths)
```

gives us the names of all the columns (and give us column numbers!)

```{r}
head(all_deaths)
```

will give us the top of the data. You can customise how many rows you see --

```{r}
head(all_deaths, 10)
```

you can use tail() to see the bottom instead

```{r}
tail(all_deaths)
```

Cool summary function -

```{r}
glimpse(all_deaths)
```

Want a little sanity? Have a look at the dataset as an actual table! Probably don't want this in code, but it might be useful in console!

```{r}
View(all_deaths)
```

Useful function in skimr:

```{r}
skim(all_deaths)
```
Gives you a pretty simple summation of the data.

Time for dplyr! select() will let you pull just the columns you want

```{r}
select(all_deaths, name, allegiances, gender, nobility, year_of_death)
```

You can also do the reverse and just nominate columns to not view.

```{r}
select(all_deaths, -name)
```

Reminder that without assigning these to an object, you're not actually doing anything besides view it.

```{r}
all_deaths_no_names <- select(all_deaths, -name)
```

will throw it into an object. If you want it to be a little more interactive, wrap the whole line in parentheses and RStudio will pop a view of the new object for you

```{r}
(all_deaths_no_names <- select(all_deaths, -name))
```

Cool shit you can do -- 

select() has a whole bunch of helper functions. Check the man page with ?

```{r}
?select
```

What if you wanted just the columns related to a book?

```{r}
select(all_deaths, contains("Book"))
```

how about related to a death?

```{r}
select(all_deaths, contains("Death"))
```

You can combine things to get useful stuff like this --

```{r}
select(all_deaths, name, contains("Death"))
```

Now for filter(). Same as select(), but for rows! Let's you return all rows that meet a condition.

Makes big use of the logical operators.

Useful one you might not know --

%in% checks for equality against *more than one* value.

Time for some examples -

```{r}
filter(all_deaths, allegiances == "Lannister")
```

```{r}
filter(all_deaths, allegiances != "Lannister")
```

Shows everyone who is or isn't a Lannister

```{r}
filter(all_deaths, allegiances %in% c("House Lannister", "Lannister"))
```

Note that this pulls more results than the initial filter! The filter is doing exact matches, so all the ones with the "House" prefix *were not included*

```{r}
filter(all_deaths, year_of_death >= 299)
```

Sidebar into a handy bonus function --

```{r}
distinct(all_deaths, allegiances)
```

will show you all the distinct values from a column. So if you're trying to glance at all the House names or something, this is a handy way to give it a glance.

Might be helpful to manipulate it a bit, so

```{r}
distinct_alleg <- distinct(all_deaths, allegiances)

arrange(distinct_alleg, allegiances)
```

Will put them all in distinct_alleg and then sort them alphabetically

```{r}
filter(all_deaths, allegiances == "Lannister" 
       & nobility == 1)
```

```{r}
filter(all_deaths, book_intro_chapter >= 5 &
         book_intro_chapter <= 10)
```

Quick note -- AND takes precedence over OR

```{r}
filter(all_deaths, year_of_death <= 299)
```


```{r}
filter(all_deaths, name %in% c("Jon Snow", "Samwell Tarly", "Daenerys Targaryen"))
```

```{r}
got_characters <- filter(all_deaths, allegiances != "Lannister" & gender == 0)
select(got_characters, name)
```
OR NEST THIS

```{r}
select(filter(all_deaths, allegiances != "Lannister" & gender == 0), name)
```

```{r}
all_deaths %>% 
  filter(allegiances != "Lannister" & gender == 0) %>%
  select(name)
```

PIPES OWN. CTRL+SHIFT+M %>% %>% %>% %>% %>% %>% %>% 

Time for arrange(). By default this is in ascending order

```{r}
all_deaths %>% 
  arrange(gender)
```

Wanna do it descending?

```{r}
all_deaths %>% 
  arrange(desc(gender))
```

You can sort my multiple columns. Sorts in priority from left to right.

```{r}
all_deaths %>% 
  arrange(book_of_death, death_chapter)
```

```{r}
all_deaths %>% 
  arrange(allegiances)
```

```{r}
all_deaths %>% 
  arrange(allegiances, book_intro_chapter)
```

```{r}
all_deaths %>% 
  arrange(desc(year_of_death))
```

