---
title: "Spatial Data"
output: html_notebook
---

# Geospatial Analysis / Visualisation in R

```{r}
library(tidyverse)
library(sf)
```

An example dataset from the sf package:

```{r}
north_carolina = st_read(system.file("shape/nc.shp", package = "sf"))
```
Spatial data looks a bit like our usual rectangular data, but with a geometry column attached.

1 row per "feature" -- geometric thing (a state, a county, a country, a motorwar, a flightpath, a city)

n columns for attributes fo the feature - (how many new born babies in that county, the population of that county, the name of the river, the id of the plane on the flightpath)

```{r}
county_geometries <- north_carolina %>% 
  st_geometry()
```

```{r}
# the first geometry
county_geometries[[1]]

plot(county_geometries[[1]])
```

Geometries are made of:

an instruction (type of geometry)
a point / series of points made using x-y co-ordinates

Types of geometries allowed: The Big 7

1. POINT((-81.47 36.23))
2. LINESTRING ((-81.47 36.23, -81.54 36.27))
3. POLYGON - first co-ord = last co-ord
 (( -81.47276 36.23436, ... , -81.47276 36.23436)))
4. MULTIPOINT
5. MULTILINESTRING
6. MULTIPOLYGON ((multiple enclosed shapes))
7. GEOMETRY COLLECTION - can have any amount of the previous 6 geometries

- we use geom_sf()

```{r}
ggplot(north_carolina) + 
  geom_sf(aes(fill = AREA), colour = "white") +
  geom_sf_text(label = north_carolina$NAME, colour = "white")
```

```{r}
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
world <- ne_countries(scale = "medium", type = "countries",
                      returnclass = "sf")

world %>% 
  head(6)
```

```{r}
ggplot(world) +
  geom_sf(aes(fill = pop_est), size = 0.1, colour = "black") +
  scale_fill_viridis_c(trans = "sqrt") +
  theme_void()
```

## Subsetting Simple Features

What if we only care about one particular country?

Can subset sf, just like filtering for particular rows, using dplyr. This time each row is a country

```{r}
faeroe_is <- world %>% 
  filter(name == "Faeroe Is.")

ggplot(faeroe_is) +
  geom_sf()
```

A really nice thing about natural earth packages. We can use ne_states to get regions in countries.

```{r}
uk <- ne_states(country = "United Kingdom", returnclass = "sf")
africa <- ne_countries(continent = "Africa", returnclass = "sf")

ggplot(uk) +
geom_sf()

ggplot(africa) +
  geom_sf(aes(fill = pop_est))

world %>% 
  filter(name == "United Kingdom") %>% 
  ggplot() +
  geom_sf()
```




## `coord_sf()` to zoom in, or change co-ordinate reference system

`expand = FALSE` means don't add a little bit more to the plotting area

```{r}
uk %>% 
  ggplot() +
  geom_sf(aes(fill = geonunit)) +
  coord_sf(xlim = c(-15, 5),
           ylim = c(49.5, 61.5),
           expand = FALSE)
```

```{r}
world %>% 
  ggplot() +
  geom_sf(aes(fill = income_grp)) +
  coord_sf(xlim = c(-102.15, -74.12),
           ylim = c(7.65, 33.97),
           expand = FALSE) +
  geom_sf_text(aes(label = name), size = 2)
```

## Leaflet

- Interactive Geospatial Visualisation

```{r}
library(leaflet)
```

Individual points (not from data)

```{r}
first_map <- leaflet() %>% 
  addTiles() %>% # adds OpenStreetMap base map by default
  addMarkers(lng = 174.768, lat = -36.852,
             popup = "The Birthplace of R")
```

```{r}
first_map
```

Getting values from data

```{r}
library(jsonlite)

colorado_data_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"

water_data_df <- fromJSON(colorado_data_url) %>%
  jsonlite::flatten(recursive = TRUE) 

# turn columns to numeric and remove NA values
water_data_df_clean <- water_data_df %>%
  mutate(across(c(amount, location.latitude, location.longitude), as.numeric)) %>%
  filter(!is.na(location.latitude))

water_data <- water_data_df_clean %>% 
  select(
    station_name, amount, starts_with("location")
  )
```

Plotting individual points on a map:

```{r}
leaflet(water_data) %>% 
  addTiles() %>% 
  addCircleM(lng = ~location.longitude, lat = ~location.latitude,
             radius = ~amount, # radius of circle
             weight = 0.1 # thickness of edge of circle
             )
```

## Other Options ..

Clustering

```{r}
leaflet(water_data) %>% 
  addTiles() %>% 
  addCircleMarkers(lng = ~location.longitude, lat = ~location.latitude,
                   clusterOptions = markerClusterOptions())
```

