---
title: "Correlation"
output: html_notebook
---
```{r}
library(tidyverse)
```


```{r}
mtcars %>% 
    select(wt, mpg) %>% 
    glimpse()
```

```{r}
mtcars %>% 
    ggplot(aes(x = wt, y = mpg)) +
    geom_point()
```

## Pearson correlation coefficient - denoted by `r`

- r has no units (it's just a number)
- it takes bounded values (betweem -1 and 1)
    - approaching r = 1 when variables are strongly positively correlated
    - approaching r = -1 when strongly negatively correlated
    - near r = 0 when not correlated
    
We can calculate r using the `cor()` function (Pearson r is the default)

```{r}
noisy_bivariate <- function(noise = 1, gradient = 1){
  x <- runif(n = 200, min = 0, max = 10)
  y <- gradient * x + 10
  y_scatter <- noise * 4 * rnorm(n = 200)
  y <- y + y_scatter
  data = tibble(x, y)

  r <- round(cor(x, y), 4)
  title <- paste(
    "noise = ", noise,
    ", gradient = ", gradient,
    ", r = ", r
  )
  
  data %>%
    ggplot(aes(x = x, y = y)) +
    geom_point() +
    xlim(0, 10) +
    ylim(min(c(min(y), 0)), max(c(max(y), 10))) +
    ggtitle(title)
  
}


noisy_bivariate(noise = 0, gradient = 2)
```

The magnitude of r: 

0 - no correlation
0.01 - 0.19 - very weak
0.2 - 0.39 - weak
0.4 - 0.59 moderate
0.6 - 0.79 - strong
0.8 - 0.99 - very strong
1 - perfect

Let's calculate the correlation coefficient for `wt` and `mpg` in mtcars

```{r}
mtcars %>% 
    summarise(cor(wt, mpg))
```

Order of variables doesn't matter

##word of warning

The correlation coefficient is a blunt instrument - we can run `cor()` on any bivariate dataset and get a number out - regardless of any real trend in the data (if any)

## The most famous phrase in statistics!

**Correlation is not causation**

Just because two variables are **correlated** does not mean that one of them **causes** the change in the other

Consider `sunglasses_sold` and the amount of `ice_cream_sold` at a beach store in one week

- there's a good chance these will be correlated
- but does this mean that the amount of sunglasses sold **causes** the amount of ice cream sold to increase?? No! We don't have a strong enough basis to make this conclusion

## The 'third variable' problem

The simplest realistic model might be that **both** these variables are themselves casually related with a third **hidden variable**

What possible third variable could cause an increase in ice cream sold and sunglasses sold?

AKA **confounding variable**

## The 'directionality' problem

Even if two variables **are** causally related, there is nothing in the statistics to indicate **directionality** of causation

e.g. we measure `amount_of_exercise` in a given week and `happiness` in a given week

and we discover positive correlation between these

we might be tempted to say that more exercise leads to happier people - but maybe happier people exercise more!

## How do we prove causation?

Ideally, with a randomised controlled experiment

These give us the basis to say whether or not our 'intervention' variable **causes** any change we find in the 'outcome' variables.