---
title: "Simple linear regresssion"
output: html_notebook
---

```{r}
library(tidyverse)
```


- understand regression by carrying out simple linear regression
- explore the topic of "least squares"
- understand how formulas are written in R
- use `lm()`

Regression - a process that lets us fit a line (or line-type shape) to a cloud of data points

An example using one dependent variable and one independent variable. One variable we want to explain, and one we are using to explain (response / predictor)

## Height and Weight

Our goal is to understand the relationship between height and weight so that we can:

- describe the data, and understand the association between our variables
- predict an individual's height given their weight (if there is a significant association)

after taking a sample of individuals heights and weights, we could perform SLR (simple linear regression) taking:

weight as our explanatory variable (x-axis)
height as our outcome variable (y-axis)

```{r}
height <- c(176, 164, 181, 168, 195, 185, 166, 180, 188, 174)
weight <- c(82, 65, 85, 76, 90, 92, 68, 83, 94, 74 )
sample <- tibble(
  weight,
  height
)

sample %>%
  ggplot(aes(x = weight, y = height)) +
  geom_point()
```

```{r}
sample %>% 
    summarise(cor = cor(weight, height))
```

y = gradient * x + intercept

y_fit = b0 + b1 * x

fit_height = b0 + b1 * weight

What is the best equation of a line fitting this data? How will we define "best"?

## A line of best fit

```{r}
line <- function(x, b0, b1){
    
    b0 + b1 * x
}
```

Let's pick some arbitrary values of b0 = 95, b1 = 1 just to see an example fitted line:

```{r}
sample %>% 
    mutate(fit_height = line(weight, 95, 1)) %>% 
    ggplot(aes(x = weight, y = height)) +
    geom_point() +
    geom_line(aes(y = fit_height), colour = "red") +
    geom_segment(aes(xend = weight, yend = fit_height))
```

How wrong are we?

How far away is each measurement from our fitted line? - A `residual`

residual = observed_height - fit_height
A definition of "best" is the "least worst". So our line of best fit, will minimise the overall residual value.

Lets add our residuals

```{r}
sample %>% 
    mutate(fit_height = line(weight, b0 = 95, b1 = 1)) %>% 
    mutate(resid = height - fit_height) %>% 
    summarise(sum_resid = sum(resid^2)) # take the square to remove "cancellation effect" of large but opposite residuals
```

```{r}
b0 = 96.8
b1 = 1

sample %>% 
    mutate(fit_height = line(weight, b0, b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    ggplot(aes(x = weight, y = height)) +
    geom_point() +
    geom_line(aes(y = fit_height), colour = "red") +
    geom_segment(aes(xend = weight, yend = fit_height))

sample %>% 
    mutate(fit_height = line(weight, b0 = b0, b1 = b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    summarise(sum_resid = sum(resid^2))
```

```{r}
b0 = 57
b1 = 1.5

sample %>% 
    mutate(fit_height = line(weight, b0, b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    ggplot(aes(x = weight, y = height)) +
    geom_point() +
    geom_line(aes(y = fit_height), colour = "red") +
    geom_segment(aes(xend = weight, yend = fit_height))

sample %>% 
    mutate(fit_height = line(weight, b0 = b0, b1 = b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    summarise(sum_resid = sum(resid^2))
```

```{r}
b0 = 77
b1 = 1.25

sample %>% 
    mutate(fit_height = line(weight, b0 = b0, b1 = b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    summarise(sum_resid = sum(resid^2))

sample %>% 
    mutate(fit_height = line(weight, b0, b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    ggplot(aes(x = weight, y = height)) +
    geom_point() +
    geom_line(aes(y = fit_height), colour = "red") +
    geom_segment(aes(xend = weight, yend = fit_height))


```

```{r}
b0 = 105
b1 = 0.9

sample %>% 
    mutate(fit_height = line(weight, b0 = b0, b1 = b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    summarise(sum_resid = sum(resid^2))

sample %>% 
    mutate(fit_height = line(weight, b0, b1)) %>% 
    mutate(resid = height - fit_height) %>% 
    ggplot(aes(x = weight, y = height)) +
    geom_point() +
    geom_line(aes(y = fit_height), colour = "red") +
    geom_segment(aes(xend = weight, yend = fit_height))


```
## Performing Linear Regression

```{r}
lm(formula = height ~ weight,
   data = sample)
```

b0 = 102.17
b1 = 0.93

If I change weight by +1, how does height change?

-> 0.9336

height = 102 + (0.9336 * weight)
diff_height = 0.9336 * diff_weight

What would be the height of someone weighs 60kg?

```{r}
102 + (0.9336 * 60)
```
## The Patsy Notation (to be continued!)

y ~ x

~: "varies with"

## Better ways of using the model

```{r}
model <- lm(formula = height ~ weight,
   data = sample)

# produces a model object; a list we can access by name
```

```{r}
model$coefficients
```
```{r}
# fit heights (for observed weights)
# effectively -- given a specific weight, what value would fit exactly on the line?
fitted(model)
```
using the `predict()` function

```{r}
predict_at <- data.frame(weight = c(60, 70, 78))

predict(model, newdata = predict_at)
```

Based on our current model of height, we expect a person weighing 60kg will measure around 160cm.

```{r}
library(modelr)
```

`{modelr}`

`add_residuals()`
`add_predictions()`

```{r}
sample %>% 
    add_predictions(model, var = "pred_height") %>% 
    add_residuals(model, var = "resid_height") %>% 
  #  summarise(resid_sum_sq = sum(resid_height^2))
    ggplot(aes(x = weight)) +
    geom_point(aes(y = height)) +
    geom_line(aes(y = pred_height), col = "blue")
```

What about predicting values not in our data (unobserved weights)

```{r}
predict_at %>% 
    add_predictions(model)

tibble (
    weight = 40:120
) %>% 
    add_predictions(model)
```

