---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
comms_data <- read_csv("data/telecom_data.csv")
```

```{r}
comms_data %>% 
  filter(is.na(monthly_charges))
```

```{r}
comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)))
```

Removing N/As - can be done globally using *drop_na()*, or by single column by specifying inside the parentheses

```{r}
comms_data %>% 
  drop_na()
```

```{r}
comms_data %>% 
  drop_na(monthly_charges)
```

Impute data -- use *coalesce()* to replace NA values directly. Can apply whatever calculation you like. *median* being used in this case

```{r}
comms_data_imputed_median <- comms_data %>% 
  mutate(monthly_charges = coalesce(monthly_charges, median(monthly_charges, na.rm = TRUE)))
comms_data_imputed_median
```

```{r}
comms_data_imputed_mean <- comms_data %>% 
  mutate(monthly_charges = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)))
comms_data_imputed_mean
```

```{r}
ggplot(comms_data_imputed_median, aes(customerID, monthly_charges)) + geom_point()
```

```{r}
ggplot(comms_data_imputed_mean, aes(customerID, monthly_charges)) + geom_point()
```

You need to watch for manually inputted values referring to some form of N/A. R *will not* understand this or identify them with *is.na()*. For instance this is not actually identifying all the rows we want --

```{r}
comms_data_imputed_median %>% 
  filter(is.na(total_charges))
```

You'll need to fix this manually.

```{r}
total_charges_replaced <- comms_data_imputed_median %>% 
  mutate(total_charges = na_if(total_charges, "na")) %>% 
  mutate(total_charges = na_if(total_charges, "N/A"))
```

6 more rows!

```{r}
total_charges_replaced %>% 
  filter(is.na(total_charges))
```

