---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
comms_data <- read_csv("data/telecom_data.csv")
```

```{r}
comms_data %>% 
  filter(is.na(monthly_charges))
```

```{r}
comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)))
```

Removing N/As - can be done globally using *drop_na()*, or by single column by specifying inside the parentheses

```{r}
comms_data %>% 
  drop_na()
```

```{r}
comms_data %>% 
  drop_na(monthly_charges)
```

Impute data -- use *coalesce()* to replace NA values directly. Can apply whatever calculation you like. *median* being used in this case

```{r}
comms_data_imputed_median <- comms_data %>% 
  mutate(monthly_charges = coalesce(monthly_charges, median(monthly_charges, na.rm = TRUE)))
comms_data_imputed_median
```

```{r}
comms_data_imputed_mean <- comms_data %>% 
  mutate(monthly_charges = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)))
comms_data_imputed_mean
```

```{r}
ggplot(comms_data_imputed_median, aes(customerID, monthly_charges)) + geom_point()
```

```{r}
ggplot(comms_data_imputed_mean, aes(customerID, monthly_charges)) + geom_point()
```

You need to watch for manually inputted values referring to some form of N/A. R *will not* understand this or identify them with *is.na()*. For instance this is not actually identifying all the rows we want --

```{r}
comms_data_imputed_median %>% 
  filter(is.na(total_charges))
```

You'll need to fix this manually.

```{r}
total_charges_replaced <- comms_data_imputed_median %>% 
  mutate(total_charges = na_if(total_charges, "na")) %>% 
  mutate(total_charges = na_if(total_charges, "N/A"))
```

6 more rows!

```{r}
total_charges_replaced %>% 
  filter(is.na(total_charges))
```

Another way to adjust total_charges here is to change the column to a numeric, since that's what it is.
This introduces NAs by "coercion", because it forces the strings out and has no choice but to replace them with NA.

```{r}
charges_numeric <- comms_data_imputed_median %>% 
  mutate(total_charges = as.numeric(total_charges))
```

```{r}
charges_numeric <- charges_numeric %>% 
  filter(!is.na(total_charges))
charges_numeric
```

Task - 5 minutes

Take a look at the payment_method column and find the different type of missing value identifiers we have, and how many we have.

Replace these with NA, and then check if you have the right amount

```{r}
charges_numeric %>% 
  group_by(payment_method) %>% 
  summarise()
```

```{r}
payment_method_replaced <- charges_numeric %>% 
  mutate(payment_method = na_if(payment_method, "--"))
payment_method_replaced %>% 
  group_by(payment_method) %>% 
  summarise()
```

If you want to clean up the NAs and actually add something

```{r}
payment_method_replaced <- charges_numeric %>% 
  mutate(payment_method = na_if(payment_method, "--")) %>% 
  mutate(payment_method = coalesce(payment_method, "unavailable"))
payment_method_replaced %>% 
  group_by(payment_method) %>% 
  summarise()
```

