---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
library(janitor)
library(modelr)
library(pROC)
library(caret)
```

```{r}
mortgage_data <- read_csv("data/mortgage_applications.csv") %>% 
  clean_names()
```

```{r}
mortgage_3pred_model <- glm(accepted ~ tu_score + employed + age,
                                        mortgage_data , family = binomial)

mortgage_data_with_3pred <- mortgage_data %>% 
  add_predictions(mortgage_3pred_model, type = "response")

mortgage_data_with_3pred
```

### 

-   Be able to plot a **receiver operating characteristic** (ROC) curve
-   and calculate the area **under the curve** (AUC) and the **Gini
    coefficient**
-   See **cross validation** applied to logistic regression
-   Understand that **expected value** calculations can be used to set
    an optimal threshold

Plot the thresholds on the axes of sensitivity vs 1-specificity TPR vs
FNR y x

```{r}
roc_obj_3pred <- mortgage_data_with_3pred %>% 
  roc(response = accepted, predictor = pred)

roc_curve <- ggroc(data = roc_obj_3pred, legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "Sensitivity (TPR)",
       x = "1 - Specificity (TNR) [FNR]")
```

```{r}
classifier_data <- tibble(
  threshold = roc_obj_3pred$thresholds,
  sensitivity = roc_obj_3pred$sensitivities,
  specificity = roc_obj_3pred$specificities
)

classifier_data
```

```{r}
ggplot(classifier_data, aes(1-specificity, sensitivity, colour = threshold)) +
  geom_line() +
  coord_equal() +
  scale_colour_gradient(low = "blue", high = "red")
```

Let's fit another classifier and add its curve to the ROC plot

```{r}
mortgage_1pred_model <- glm(accepted ~ tu_score,
                                        mortgage_data , family = binomial)

mortgage_data_with_1pred <- mortgage_data %>% 
  add_predictions(mortgage_1pred_model, type = "response")

roc_obj_1pred <- mortgage_data_with_1pred %>% 
  roc(response = accepted, predictor = pred)

roc_curve1_3 <- ggroc(data = list(roc_obj_1pred, roc_obj_3pred), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "Sensitivity (TPR)",
       x = "1 - Specificity (TNR) [FNR]")

mortgage_1pred_age <- glm(accepted ~ age,
                                        mortgage_data , family = binomial)

mortgage_data_with_age <- mortgage_data %>% 
  add_predictions(mortgage_1pred_age, type = "response")

roc_obj_age <- mortgage_data_with_age %>% 
  roc(response = accepted, predictor = pred)

roc_curve1_3_age <- ggroc(data = list("TU Score" = roc_obj_1pred, "Age" = roc_obj_age, "3 Preds" = roc_obj_3pred), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "Sensitivity (TPR)",
       x = "1 - Specificity (TNR) [FPR]")
```

### Area Under the Curve (AUC)

```{r}
auc(roc_obj_3pred)
auc(roc_obj_1pred)
```

### Gini Coefficient

Transformation of the AUC such that the values are constrained between
-1 and 1

$$
Gini = 2 \times AUC - 1
$$

```{r}
gini3 <- 2 * auc(roc_obj_3pred) - 1
gini1 <- 2 * auc(roc_obj_1pred) - 1
```

```{r}
mortgage_data <- mortgage_data %>% 
  mutate(across(where(is_logical), ~as_factor(tolower(.))))
```

`trainControl()` K-fold Cross-Validation How many K? 5

```{r}
mod_form <- accepted ~ tu_score + employed + age

train_control <- trainControl(method = "repeatedcv",
                              number = 5,
                              repeats = 100,
                              savePredictions = TRUE,
                              classProbs = TRUE,
                              summaryFunction = twoClassSummary)

model <- train(mod_form,
               data = mortgage_data,
               trControl = train_control,
               method = "glm",
               family = binomial(link = "logit"))
```

**expected value**

Is the sum-product of `value * prob`

multiply each value by its probability, then sum them up to get the
expected value

### Scenario

Company referall campaign. Target mortage applicants. Focus our campaign
on those that are most likely to be accepted

Contractor gets paid every time they recommend *accept* and the bank
*does* accept.

*true postitive* (highly eligible applicant correctly identified as such)
- profit the company a referral fee of `£25`
- true positive profit *TPP*

*true negative* (a low eligibility applicant corretly identified as such)
- earns no profit, as no approach is made.
- true negative profit *TNP*

*false positive* (a low eligibility applicant misidentified as highly eligible)
- "profits" the company `-£5` due to contact and administration costs
- false positive profit *FPP*

*false negative* (a highly eligible applicant misidentified as low eligibility)
- earns no profit, as no approach is made
- false negative profit *FNP*

```{r}
classifier_data <- classifier_data %>% 
  rename(tpr = sensitivity,
         tnr = specificity) %>% 
  mutate(fpr = 1 - tnr,
         fnr = 1 - tpr)
```

```{r}
sum(mortgage_data$accepted == "true") / 1000

```
```{r}
prob_pos <- mean(mortgage_data$accepted == "true")
prob_neg <- mean(mortgage_data$accepted == "false")

# profits
tpp <- 20
tnp <- 0
fpp <- -5
fnp <- 0

classifier_data <- classifier_data %>% 
  mutate(exp_profit_per_applicant = prob_pos * (tpr * tpp + fnr * fnp) +
                                    prob_neg * (tnr * tnp + fpr * fpp))

classifier_data %>% 
  arrange(-exp_profit_per_applicant)

classifier_data %>% 
  arrange(fpr)
```

```{r}
classifier_data %>% 
  ggplot(aes(threshold, exp_profit_per_applicant)) +
  geom_line()
```

