---
title: "Visual Encoding"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
total_sales <- read_csv("data/total_sales.csv")
```

```{r}
total_sales %>% 
  ggplot(aes(branch, sales)) +
  geom_col() +
  coord_flip()
```

```{r}
total_sales %>% 
  ggplot(aes(x = reorder(branch, sales), y = sales)) +
  geom_col() +
  coord_flip()
```

```{r}
total_sales %>% 
  ggplot(aes(branch, sales, fill = branch == "Leeds")) +
  geom_col() +
  scale_fill_manual(values = c("grey40", "orange"), guide = "none") +
  coord_flip()
```

```{r}
annotate <- total_sales %>% 
  mutate(label_text = ifelse(branch == "Leeds",
                             "Leeds had 715k sales",
                             NA))

annotate %>% 
  ggplot(aes(branch, sales,
             fill = branch == "Leeds")) +
  geom_col() +
  scale_fill_manual(values = c("grey40", "orange"),
                    guide = "none") +
  coord_flip() +
  geom_text(aes(label = label_text),
            colour = "orange", size = 4, hjust = 0)
```

This graph is pretty cluttered and it’s not clear what we are trying to show.
Rewrite the graph above in a more compelling manner. Consider how you could emphasise which branches have been most successful in the month of April compared to January.

```{r}
monthly_sales <- read_csv("data/monthly_sales.csv") %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April")))

head(monthly_sales)
ggplot(monthly_sales) +
  aes(x = reorder(branch, -difference_from_jan), y = sales, fill = month) +
  geom_col(position = "dodge") +
  coord_flip()
```

```{r}
monthly_sales %>% 
  ggplot(aes(x = month, y = sales, group = branch, col = branch)) +
  geom_line()
```

```{r}
ggplot(monthly_sales) +
  aes(x = month, y = sales, group = branch, colour = branch) +
  geom_line() 
```


A senior manager suspects the proportion of sales arising from Northern offices (Edinburgh, Glasgow, Leeds, Manchester) are growing relative to southern offices. Produce a visualisation that speaks to this.

```{r}
monthly_sales %>% 
  mutate(region = if_else(branch %in% c("Edinburgh", "Glasgow", "Leeds", "Manchester"), "North", "South")) %>% 
  ggplot(aes(x = month, y = sales, group = branch, colour = branch)) +
  geom_line() +
  facet_wrap(~ region, scales = "free_y")
```
Task
The dataset got_ratings contains the episode ratings for the TV series Game of Thrones. Produce a graph.
The viewer should:
be able to understand the chronology of the ratings
be able to attribute each rating to a particular season
be able to view some indicator of the perceived quality of each season overall
notice the perceived decline of the series in the final season
You do not need to use the data on number of reviews.

```{r}
got_ratings <- read_csv("data/got_ratings.csv")
```

```{r}
got_ratings %>% 
  mutate(show_avg = mean(average_rating)) %>% 
  group_by(series) %>% 
  mutate(series_avg = mean(average_rating)) %>% 
  ungroup() %>% 
  ggplot(aes(x = episode, y = average_rating, colour = average_rating)) +
  geom_line() +
  geom_line(aes(episode, y = series_avg), colour = "red") +
  geom_line(aes(episode, y = show_avg), colour = "orange") +
  scale_x_continuous(breaks = 1:10) +
  facet_wrap(~ series, scales = "free_y")
```

