---
title: "Multiple Regression"
output: html_notebook
---

```{r}
library(mosaic) # for model plots
library(mosaicData) # for our RailTrail data
library(janitor)
library(tidyverse)
library(GGally) # for pair plot
library(ggfortify) # for model diagnostics
```

# Elements of Multiple Regression

- multiple terms in regression equation
- interpretation of multiple regression
- categorical predictors (+ interpretation)
- interactions of predictors

Simple Linear Regression:

height ~ weight

Multiple Linear Regression:

volume ~ avgtemp + cloudiness + weekend

bill_length ~ species + bill_depth

ice_cream ~ temperature + sunny_day + location_glasgow

## setup - understanding the data

```{r}
RailTrail
```

## Data Preparation (Variable Engineering)

- dayType isn't snakecase
- cloudcover isn't one word
- some columns contain the same information (duplicate info)
- the seasonal columns are encoded as integers even though that are logical/boolean type (TRUE / FALSE) [there's also no winter in this data]

```{r}
railtrail_clean <- RailTrail %>% 
    clean_names() %>% 
    rename(cloudage = cloudcover) %>% 
    mutate(across(spring:fall, as.logical))
```

```{r}
head(railtrail_clean)
```

`alias()` can detect perfect collinearity (this is when a variab;e can be constructed by another / a combination of other variables in the data)

```{r}
# volume ~ (. means every other variable)
alias(volume ~ ., data = railtrail_clean)
```

From our alias table we see that 
    - avgtemp is a combination of lowtel and hightemp
    - fall is work-outable from spring and summer
    - day_type is work-outable from weekday
    
```{r}
railtrail_trim <- railtrail_clean %>% 
    select(-c(lowtemp, hightemp, fall, day_type))
```
    
```{r}
# check there are no aliassed variables
alias(volume ~ ., data = railtrail_trim)
```

## a pairs plot

- a tool for examining the relationship between all pairs of variables

```{r}
ggpairs(railtrail_trim, progress = FALSE)
```

What are we looking for?

Significant relationships between:

- our outcome and any predictor
- any predictor and another predictor

Variables we might want to try modelling:

- cloudage
- avgtemp
- precip
- Summer
- weekday

## Starting point: simple linear regression

volume ~ avgtemp

cor = 0.427

volume = b0 + b1 * avgtemp

```{r}
model <- lm(volume ~ avgtemp, data = railtrail_trim)

autoplot(model)
```

```{r}
summary(model)
```
Model results:

avgtemp is a significant term in our model
avgtemp only provides 0.18 (18%) proportion of variance explained in volume

rse = 115.9

```{r}
ggplot(railtrail_trim, aes(avgtemp, volume)) +
    geom_point()
```

volume = b0 + b1 * avgtemp

b0 IS INTERCEPT COEFFICIENT ESTIMATE
B1 IS AVGTEMP COEFFICIENT ESTIMATE

volume = 99.602 + 4.802 * 60 = 387.7

# Building a (hopefully) better model: multiple regression

Adding a categorical variable to a linear regression

Task - 2 mins
Try plotting an appropriate visualisation to determine whether user volume is associated with the weekday predictor.

```{r}
railtrail_trim %>% 
    ggplot(aes(x = weekday, y = volume)) +
    geom_col()
```
Can also grav the correlation coefficient:

```{r}
railtrail_trim %>% 
    summarise(cor = cor(weekday, volume))
```
-0.2932675

a weak negative correlation (expect more people using the trail on weekends compared to weekdays)

so, to add weekday to the model --

volume = b0 + b1 * avgtemp + b2 * weekday

```{r}
model_2 <- lm(
    volume ~ avgtemp + weekday,
    data = railtrail_trim
)

autoplot(model_2)
summary(model_2)
```

plot 1: no clear pattern to residuals
plot 2: residuals appear normally distributed
plot 3: again, some evidence of heteroskedasticity but variance appears reasonably uniform.

b0 = 163.022  
b1 = 4.541
b2 = -70.320

how many users if avgtemp is 65 on a saturday?

volume = b0 + b1 * avgtemp + b2 * weekday

volume = 163.02 + 4.54 * 65 + -70.32 * 0 = 458 [we'd expect 70 fewer people on a weekday]

```{r}
plotModel(model_2)
```

We call this a __parallel slopes model__, whether weekday is "on" or "off" the slope of the line does not change.

Task - 5 mins

Try adding the summer categorical predictor to the existing model with avgtemp and weekday.

How many lines do you expect to see in this model?
Is this a parallel slopes model? [Hint try plotModel() on the model object]
Is the addition of this predictor justified [Hint what is the p-value of b_summer]?

```{r}
model_3 <- lm(
    volume ~ avgtemp + weekday + summer,
    data = railtrail_trim
)

autoplot(model_3)
summary(model_3)
plotModel(model_3)
```

As summer does not seem to be a statistically significant term in our model (perhaps due to high variation being explained from avgtemp already), we won't add it as a term

Current model:

volume ~ avgtemp + weekday

- interactions: terms involving two or more predictors

Does the volume of users on the trail depend on whether or not it's a weekday and how does temperature factor into this?

```{r}
railtrail_trim %>% 
    ggplot(aes(x = avgtemp, y = volume, colour = weekday)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)
```
What does this say?

The two lines have different gradients. On a weekday, increasing the avgtemp results in a larger increase in users.

How do we add this into our model?

- add an interaction

volume = b0 + b1 * avgtemp + b2 * weekday + b3 * avgtemp * weekday

equation of the two lines:

when weekday = 0
volume = b0 + b1 * avgtemp

when weekday = 1 
volume = b0 + b1 * avgtemp + b2 + b3 * avgtemp

or in it's "simplest" form
volume = (b0 + b2) + (b1 + b3) * avgtemp

formula is volume varies with average temp and weekday and the interaction of avg temp and weekday

```{r}
model_interaction <- lm(
    formula = volume ~ avgtemp + weekday + avgtemp:weekday,
    data = railtrail_trim
)
```

```{r}
summary(model_interaction)
```
```{r}
autoplot(model_interaction)
```

While the diagnostic plots look fine (ish - still a bit of heteroskedasticity in Scale-Location), the p-vlaues suggest the interaction is not statistically significant.

```{r}
model_3terms <- lm(
    volume ~ avgtemp + weekday + cloudage,
    data = railtrail_trim
)
```

```{r}
summary(model_3terms)
```
Interpreting this

volume = b0 + b1 * avgtemp + b2 * weekday + b3 * cloudage

y = ax + bz + c 

how many users on a saturday if 5 cloudage and 55F

200.1 + 5.2 * 55 + -47.9 * 0 + -16 * 4

200.1 + 5.2 * 55 + -16 * 5

406
422
411