---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

New Week, new topic! Joins is up first.

Database relationships and keys, briefly:

### Relationships

one-to-one
many-to-many
one-to-many

*Keys* are identifiers that are used to make rows uniquely identifiable.

Primary keys are the unique ID for a table, foreign keys are primary keys from a different table referenced somewhere else.

Joins are done using `dplyr`

### Mutating joins

`mutating-joins`

`inner_join()`
  Will return all rows from table x that have a **match** in table B
  
`left_join()`
  Includes all rows in table x
  Will also **add** columns from table y
  
`right_join()`
  Includes all rows in table x
  Will also **add** columns from table y

`full_join()`
  Includes all rows in table x and table y

### Filtering joins

`filter-joins`

Filtering joins **filter rows** from table x based on the presence or absence of matches in table y

`semi_join()` returns **only** rows from table x where there are at least one match in table y

`anti_join()` returns only rows from table X where there are **no** matches in table Y

### Binding functions

No concept of "matching" 

`bind_rows()`
  Adds rows to a table
  
`bind_cols()`
  Adds columns to a table

```{r}
count()
```

```{r}
customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)


orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)


customers
orders
```

Customers --
Primary key = customer_ID

Orders --
Primary key = order_number
Foreign key = customer_ID

### `inner_join()`

Consider    `customers` as Table X
            `orders` as Table y
            
```{r}
customers %>% 
  inner_join(orders, by = "customer_ID")
```

### `left_join()`

Includes all rows in x

```{r}
left_join(customers, orders)
left_join(orders, customers)
```

It appends the columns required from table y into table x.

`right_join` works the same way, but does it the other way round. Functionally the difference is about column layout in the new output -> they're creating same dataframe otherwise.

```{r}
right_join(customers, orders)
right_join(orders, customers)
```
### `full_join`

```{r}
full_join(customers, orders, by = "customer_ID")
```

Both combined adds everything, including rows where no relationship can be found. 

### More on the "by ="

Often you'll need to identify which columns are the common relationship, for example primary key to foreign key

```{r}
customers %>% 
  inner_join(orders, by = "customer_ID")

customers %>% 
  rename(this_is_the_primary_key = customer_ID) %>% 
  inner_join(orders, c("this_is_the_primary_key" = "customer_ID"))
```

```{r}
customers2 <- customers %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))

orders2 <- orders %>%
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "cash"))

customers2
orders2

# join by two columns
inner_join(customers2, orders2, by = c("customer_ID", "payment_type"))
```

Note that because `customers2` has a payment_type column, the join matches across that and eliminates John Smith's payment that doesn't match up

### filtering joins

```{r}
semi_join(customers, orders, "customer_ID")
```

Returns all rows from `customers` when their customer_ID occurs at least once in `orders`.
Basically it's an `inner_join` with no dupes?

```{r}
anti_join(customers, orders, "customer_ID")
```

```{r}
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 
```

If you run bind_cols() on your customers and statuses tables, how many rows will be returned?
Use the bind_cols() function to merge customers and statuses.

```{r}
bind_cols(customers, statuses)
```

```{r}
past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  shipping_address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)

bind_rows(customers, past_customers)
```

