---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

New Week, new topic! Joins is up first.

Database relationships and keys, briefly:

### Relationships

one-to-one
many-to-many
one-to-many

*Keys* are identifiers that are used to make rows uniquely identifiable.

Primary keys are the unique ID for a table, foreign keys are primary keys from a different table referenced somewhere else.

Joins are done using `dplyr`

### Mutating joins

`mutating-joins`

`inner_join()`
  Will return all rows from table x that have a **match** in table B
  
`left_join()`
  Includes all rows in table x
  Will also **add** columns from table y
  
`right_join()`
  Includes all rows in table x
  Will also **add** columns from table y

`full_join()`
  Includes all rows in table x and table y

### Filtering joins

`filter-joins`

Filtering joins **filter rows** from table x based on the presence or absence of matches in table y

`semi_join()` returns **only** rows from table x where there are at least one match in table y

`anti_join()` returns only rows from table X where there are **no** matches in table Y

### Binding functions

No concept of "matching" 

`bind_rows()`
  Adds rows to a table
  
`bind_cols()`
  Adds columns to a table

```{r}
count()
```

```{r}
customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)


orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)


customers
orders
```

Customers --
Primary key = customer_ID

Orders --
Primary key = order_number
Foreign key = customer_ID

### `inner_join()`

Consider    `customers` as Table X
            `orders` as Table y
            
```{r}
customers %>% 
  inner_join(orders, by = "customer_ID")
```

### `left_join()`

Includes all rows in x

```{r}
left_join(customers, orders)
left_join(orders, customers)
```

It appends the columns required from table y into table x.

`right_join` works the same way, but does it the other way round. Functionally the difference is about column layout in the new output -> they're creating same dataframe otherwise.

```{r}
right_join(customers, orders)
right_join(orders, customers)
```
### `full_join`

```{r}
full_join(customers, orders, by = "customer_ID")
```

Both combined adds everything, including rows where no relationship can be found. 

### More on the "by ="

Often you'll need to identify which columns are the common relationship, for example primary key to foreign key

```{r}
customers %>% 
  inner_join(orders, by = "customer_ID")

customers %>% 
  rename(this_is_the_primary_key = customer_ID) %>% 
  inner_join(orders, c("this_is_the_primary_key" = "customer_ID"))
```

```{r}
customers2 <- customers %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))

orders2 <- orders %>%
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "cash"))

customers2
orders2

# join by two columns
inner_join(customers2, orders2, by = c("customer_ID", "payment_type"))
```

Note that because `customers2` has a payment_type column, the join matches across that and eliminates John Smith's payment that doesn't match up

### filtering joins

```{r}
semi_join(customers, orders, "customer_ID")
```

Returns all rows from `customers` when their customer_ID occurs at least once in `orders`.
Basically it's an `inner_join` with no dupes?

```{r}
anti_join(customers, orders, "customer_ID")
```

```{r}
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 
```

If you run bind_cols() on your customers and statuses tables, how many rows will be returned?
Use the bind_cols() function to merge customers and statuses.

```{r}
bind_cols(customers, statuses)
```

```{r}
past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  shipping_address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)

bind_rows(customers, past_customers)
```

### Tidy Data

Fundamentals of tidy data -

1. Each variable should have its own column
2. Each observation should have its own row
3. Each value should have its own cell

Four main functions in `tidyr`

`pivot_longer()`
    makes data into a "longer" format

`pivot_wider()`
    makes data into a "wider" format
    
`separate()` 
    splits column data
  
`unite()`
    to join columns
    
```{r}
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
  hr_am = c(68,72,68), 
  hr_pm = c(58,89,52))

subject_hr

subject_hr %>% 
  pivot_longer(cols = c(hr_am, hr_pm), 
               names_to = "measurement_time",
               values_to = "heartrate")
```
```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))

income

tidy_income <- income %>% 
  pivot_longer(-religion, 
               names_to = "income",
               values_to = "count")

tidy_income
```

```{r}
messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
  order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
  order_amount = c(58.00,125.00,75.25))

messy_orders

tidy_orders <- messy_orders %>% 
  separate(order_info, c("order_number", "date"),
           sep = "/")
tidy_orders
```

```{r}
hospital_visit <- read_csv("data/hospitals93to98.csv")
```

```{r}
hospital_visit

hospital_visits_longer <- hospital_visit %>% 
  pivot_longer(cols = starts_with("FY"),
               names_to = "year",
               values_to = "count")
hospital_visits_longer

hospital_visits_wider <- hospital_visits_longer %>% 
  pivot_wider(names_from = Field,
              values_from = count)
hospital_visits_wider

hospital_coded <- hospital_visits_wider %>% 
  separate(IcdChapter, c("code", "description"),
           sep="\\.")
hospital_coded
```

Wouldn't normally use `unite` on this data, but included for completeness!

```{r}
hospital_unite <- hospital_coded %>% 
  unite(disease, c("code", "description"), sep = ":")

head(hospital_unite)
```

