---
title: "Tidy Data Mini-Lab"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

**Duration: 30 minutes** 

Now it's your turn to practice the verbs on a new dataset. Complete the following task:

# MVP

**Task 1.**  
Load in the `country_data.csv` file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

<br>

**Task 2.**  
Do the appropriate pivoting of your columns using R  

<br>

**Task 3.**  
Separate the `country` variable so you have the name in one column and the number in another.   

<br>

**Task 4.**  
Unite the `code_letter` and `code_numb` variables so that they are both in one column, with no spaces separating them.   

# Answer


```{r}
country_data <- read_csv("data/country_data.csv")
```


## Task 1

Years, Type

## Task 2

```{r}
country_data_longer <- country_data %>% 
  pivot_longer(cols = starts_with("Y"),
               names_to = "year",
               values_to = "count")
country_data_longer

country_data_wider <- country_data_longer %>% 
    pivot_wider(names_from = type,
                values_from = count)
country_data_wider
```

# Task 3

```{r}
country_data_coded <- country_data_wider %>% 
  separate(country, c("id", "name"),
           sep="\\.")
country_data_coded
```
# Task 4

 
```{r}
country_united <- country_data_coded %>% 
  unite(code, c("code_letter", "code_numb"), sep = "")
country_united
```
