---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

Recoding! dplyr functions!

`distinct()` is a dplyr function that takes in a dataset and a list of columns, and 
returns the unique values in that column

```{r}
students_big %>% 
  distinct(region)
```

`count()` works similarly, but returns the number of occurences of each distinct
value

```{r}
students_big %>% 
  count(region)
```

`count()` is cool, because you can also apply it to multiple columns to get more
distinct breakdowns of the data.

`distinct()` also does this!

```{r}
students_big %>% 
  count(region, gender)
```

Recoding! Involves changing ("recoding") the values in column. Primarily to make data uniform
for easier interpretation.

Methods:
`if_else()`
`recode()`
`case_when()`

## `if_else()`

```{r}
letters <- c("a","B","c")

if_else(letters == "B", "b", "e")
```

if_else requires an else! if want it to do nothing, feed the vector in

```{r}
letters <- c("a","B","c")

if_else(letters == "B", "b", letters)
```

```{r}
students_big %>% 
  mutate(gender = if_else(gender == "M", "Male", "Female")) %>% 
  count(gender)
```

Task - 5 minutes

1. Change the column handed to to "Right" and "Left" instead or "R" and "L".
2. In favourite_physical_activity change "Football/Soccer" into "Football" and leave all other activities the same.

```{r}
students_big %>% 
  mutate(handed = if_else(handed == "R", "Right", handed)) %>% 
  mutate(handed = if_else(handed == "L", "Left", handed)) %>% 
  count(handed)
```

```{r}
students_big %>% 
  mutate(favourite_physical_activity = 
           if_else(favourite_physical_activity == "Football/Soccer", "Football",
                   favourite_physical_activity)) %>% 
  count(favourite_physical_activity)
```

## `recode`

how it works -

recode(column/vector, "old_value" = "new_value", 
        "another_old_value" = "another_new_value", ...,
        .default = NULL, .missing = NULL)
        
```{r}
students_big %>% 
  mutate(handed = recode(handed, "L" = "Left", "R" = "Right", "B" = "Both")) %>% 
  count(handed)
```

```{r}
students_big %>% 
  mutate(travel_to_school = recode(travel_to_school,
                                   "Car" = "car",
                                   "Bus" = "bus",
                                   "Walk" = "walk",
                                   .default = "other"))
```


```{r}
data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")
```

Using the data below, do the following:
Recode the sex column to be “Male” and “Female”
Recode the compl column to be T = “True”, F = “False”, and NA otherwise

```{r}
data_recode <- data %>% 
  mutate(sex = recode(sex,
                      "M" = "Male",
                      "F" = "Female")) %>% 
  mutate(compl = recode(compl,
                        "T" = "True",
                        "F" = "False",
                        .default = NA_character_))

data_recode
```

# `case_when()`

Oh boy. `~` used for conditional tests. NOTE `case_when` stops running through
the conditions when it finds the first one. *TRUE* simply enforces a case at the
end of the conditionals, effectively acting as a default replacement if no condition met.

```{r}
case_when(
  letters == "a" ~ "A"
  letters == "b" ~ "B"
  letters == "c" ~ "C",
  TRUE ~ letters)
```

```{r}
students_big %>% 
  mutate(education_level = case_when(
    ageyears > 16 ~ "sixth form",
    ageyears > 12 ~ "high school",
    ageyears <= 12 ~ "primary school")) %>% 
  count(education_level)
```

Task - 5 minutes
Using case_when, make a new column estimated_travel_time which is:
Equal to 60 when your region is London.
If you aren’t from London, then:
10 when your travel_to_school is “Walk”.
20 when your travel_to_school is “Car”.
For everyone one else then it should equal to 30.

```{r}
students_big %>% 
  mutate(estimated_travel_time = case_when(
    region == "London" ~ 60,
    travel_to_school == "Walk" ~ 10,
    travel_to_school == "Car" ~ 20,
    TRUE ~ 30)) %>% 
```

