---
title: "Week 2 Day 3"
output: html_notebook
---

# Lists

```{r}
grocery_list <- list("Bread","Butter","Milk","Chicken","Cat Food")
```

```{r}
x <-  list(1,"a",TRUE,c(1,2,3))

x[[2]]
```

Name your lists

```{r}
named_list <- list(a = "I like programming in R", b = 1:10)

named_list$a
named_list$b

names(named_list)

named_list
```

Nested lists

```{r}
student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info[[1]]




```

```{r}
student_info$Agnes
```

```{r}
student_info$Agnes$height
```

```{r}
student_info[[1]][[1]]
```

```{r}
demo_list <- list(1,"uzair","codeclan",c(1,2,3,4))

demo_list[2] <- "UZAIR"

demo_list[4] <- NULL
demo_list
```

**Task - 5 minutes**
Make a named list that contains a numeric vector and a character vector.
Access the second element of your list using double square brackets [[]]. What is its class?
Access the second element of your list using single square brackets []. What is its class?
Access the second element of your list by name. What is its class?

```{r}
task_list <- list(nums = c(1:26), alphas = c(letters))

class(task_list[[2]])
class(task_list[2])

class(task_list$alphas)
```

# Matrices

```{r}
matrix(1:4,nrow=2,ncol=2)
```

```{r}
matrix(1:4,nrow=2,ncol=2,byrow=TRUE)
```

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
m
```

```{r}
m[1,]
m[,1]
m[1,1]
m[2,3]
m[,c(1,3)]
```
```{r}
m * 3

m - 2 
```

```{r}
rowSums(m)
colSums(m)
rowMeans(m)
colMeans(m)
```

**Task - 5 minutes**
Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
Find the element in the 2nd row, 3rd column.
Check that all the columns are identical by checking that their sums are the same.

```{r}
task_matrix <- matrix(1:3, nrow=3, ncol=4)

task_matrix[2,3]
colSums(task_matrix)
```

# Control Flow

## if

```{r}
if ( 1 %in% c(1,4,6)){
  print("yes")
}
```
```{r}
if (5 != 3){
  print("Yes")}
```

```{r}
x <- 4
y <- 7

if (x > y){
  print("Yes")
}
```

**Task - 5 minutes**

Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x <- 4

if (x %% 2 == 0){
  print("Yes that divides by 2")
}
```

```{r}
z <- 33

if (z %% 2 != 0){
  print("No this is not a prime number")
}
if (z %% 2 == 0){
  print("Yes this is a prime number")
}
```

```{r}
if ( 10 != 5){
  if (3 > 1){
    print("Hello World!")
  }
  print("Goodbye!")
}
```

```{r}
if (10 != 5){
  if (2 <= 0){
    print("Hello World!")
  }
  print("Goodbye!")
}
```
**Task - 10 minutes**
Here’s an example of a nested if statement. Predict what the code is going to print out for the following combinations:
x is 10 and y is 3.
x is 10 and y is 2.
x is 2 and y is 10.

```{r}

if (x > y) {
  print("Start of outer if block")

  if (x %% y == 0) {
    print("Inner if block 1")
  }

  if (y %in% c(1:5)) {
    print("Inner if block 2")
  }

  print("End of outer if block")
}

print("After if statement")
```

1. "Start of outer if block", "Inner if block 1", "End of outer if block", "After if statement"
2. "Start of outer if block", "Inner if block 1", "Inner if block 2", "End of outer if block", "After if statement"
3. "After if statement"

## if else

```{r}
weather = "Windy"
if (weather == "Windy"){
  print("weather is windy")
} else{
  print("weather is not windy today lets go shopping")
}
```
```{r}
if (4>5){
  print("Yes")
} else{
  print("No")
}
```

```{r}
if (3 %in% 4:6){
  print("3 is in range")
} else {
  print("its not in range")
}
```

**Task - 5 minutes**

Edit your code from the first exercise so that it prints ‘Yes that divides’ when x is divisible by 2, and prints ‘No that does not divide’ otherwise.

```{r}
x <- 4

if (x %% 2 == 0){
  print("Yes that divides by 2")
}else {print("No that does not divide")
    }
```

**Task - 10 minutes**
Here’s an example of a nested if-else statement. Predict what it’s going to print out for the following conditions:.
1. When x is 5 and y is 10
2. When x is 6 and y is 12
3. When x is 6 and y is 13

```{r}
x <- 6
y <- 13

if (x %in% c(3, 4, 5)){
  
  print("x is in our list")
  
  if (x == 5){
    
    print("x is exactly 5")
    
  }
  
} else {
  
  print("Start of else statement")
  
  if (y %% x == 0){
    
    print("x divides y")
  
  } else{
    
    print("x does not divide Y")
  }
  
  print ("End of else statement")
  
}
```

1. "x is in our list", "x is exactly 5"
2. "Start of else statement", "x divides y", "End of else statement"
3. "Start of else statement", "x does not divide y", "End of else statement"

## for

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (day in weekdays) {
    phrase <- paste("The day is", day)
    print(phrase)
    # or could do in one line as
    # print(paste("The day is", day))
} 
```

**Task - 10 mins**

Change the code below to use a for loop:
```{r}
print(1)
print(2)
print(3)
print(4)
print(5)
```

```{r}
for(i in 1:5){
  print(i)
}
```
## while

```{r}
start_value <- 1
stop_value <- 10

while (start_value < stop_value){
  start_value <- start_value + 1
  phrase = paste("start value is", start_value)
  print(phrase)
}
```

# Alternative indexing

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (i in 1:length(weekdays)){
  phrase <- paste("The day is", weekdays[i])
  print(phrase)
}
```

```{r}
for (i in 1:length(weekdays)){
  weekdays[i] <- paste0(weekdays[i], "s")
}
```

# Nested loops

```{r}
my_matrix <- replicate(5, rnorm(5))

```

```{r}
for (row_index in 1:dim(my_matrix)[1]){
  for(col_index in 1:dim(my_matrix)[2]){
    print(my_matrix[row_index, col_index])
  }
}
```
**Task**

Create a nested for loop and if statement combination which loops through the weekdays and prints out the phrase ‘Last weekday!’ once it gets to Friday.

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
for (i in seq_along(weekdays)){
  if(weekdays[i] == "Friday"){
    phrase <- paste0("It's ",weekdays[i],"! Last weekday! Freedom!")
    print(phrase)
  }
}
```

# Functions

In R, functions return the last output by default. The CodeClan style guide suggests to not use `return()` unless you're doing something like if statements and have multiple different possible return values.

```{r}
function_name <- function(argument_1, argument_2 = value){

     # Do stuff
      
}
```

```{r}
multiply_by_5 <- function(x){
  x*5
}

multiply_by_5(14)
```

```{r}
multiply_by <- function(x, y){
  x * y
}

multiply_by(3, 12)
```

```{r}
multiply_by_5_def <- function(x, y = 5){
  x * y
}

multiply_by_5_def(4, 2)
```

```{r}
triangle_f <- function(x){
  (x*(x + 1))/2
}

```
**Task - 5 minutes**
Write a function make_price() that takes as input any whole number, and returns it written as a price (i.e. with a ‘£’ in front of the number). For example 5 would become £5, and 123 would become £123. Try out your function a few times to test that it works.
Hint
Use the str_c() function

```{r}
make_price <- function(x){
  if(is.numeric(x) == TRUE){  
    str_c("£", x)}else{
      print("that isn't a number")
    }
}

make_price("word")
```

