---
title: "Week 2 Day 3"
output: html_notebook
---

# Lists

```{r}
grocery_list <- list("Bread","Butter","Milk","Chicken","Cat Food")
```

```{r}
x <-  list(1,"a",TRUE,c(1,2,3))

x[[2]]
```

Name your lists

```{r}
named_list <- list(a = "I like programming in R", b = 1:10)

named_list$a
named_list$b

names(named_list)

named_list
```

Nested lists

```{r}
student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info[[1]]




```

```{r}
student_info$Agnes
```

```{r}
student_info$Agnes$height
```

```{r}
student_info[[1]][[1]]
```

```{r}
demo_list <- list(1,"uzair","codeclan",c(1,2,3,4))

demo_list[2] <- "UZAIR"

demo_list[4] <- NULL
demo_list
```

**Task - 5 minutes**
Make a named list that contains a numeric vector and a character vector.
Access the second element of your list using double square brackets [[]]. What is its class?
Access the second element of your list using single square brackets []. What is its class?
Access the second element of your list by name. What is its class?

```{r}
task_list <- list(nums = c(1:26), alphas = c(letters))

class(task_list[[2]])
class(task_list[2])

class(task_list$alphas)
```

# Matrices

```{r}
matrix(1:4,nrow=2,ncol=2)
```

```{r}
matrix(1:4,nrow=2,ncol=2,byrow=TRUE)
```

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
m
```

```{r}
m[1,]
m[,1]
m[1,1]
m[2,3]
m[,c(1,3)]
```
```{r}
m * 3

m - 2 
```

```{r}
rowSums(m)
colSums(m)
rowMeans(m)
colMeans(m)
```

**Task - 5 minutes**
Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
Find the element in the 2nd row, 3rd column.
Check that all the columns are identical by checking that their sums are the same.

```{r}
task_matrix <- matrix(1:3, nrow=3, ncol=4)

task_matrix[2,3]
colSums(task_matrix)
```

# Control Flow

```{r}
if ( 1 %in% c(1,4,6)){
  print("yes")
}
```
```{r}
if (5 != 3){
  print("Yes")}
```

```{r}
x <- 4
y <- 7

if (x > y){
  print("Yes")
}
```

**Task - 5 minutes**

Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x <- 4

if (x %% 2 == 0){
  print("Yes that divides by 2")
}
```

```{r}
z <- 33

if (z %% 2 != 0){
  print("No this is not a prime number")
}
if (z %% 2 == 0){
  print("Yes this is a prime number")
}
```

```{r}
if ( 10 != 5){
  if (3 > 1){
    print("Hello World!")
  }
  print("Goodbye!")
}
```

```{r}
if (10 != 5){
  if (2 <= 0){
    print("Hello World!")
  }
  print("Goodbye!")
}
```



