---
title: "Manual Model Development - Lab"
output: html_notebook
---

```{r}
library(car)
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
```

```{r}
prestige <- Prestige
```

```{r}
glimpse(prestige)

colSums(is.na(prestige))
```

```{r}
prestige_trim <- prestige %>% 
    drop_na() %>% 
    select(-census)
```

```{r message=FALSE}
ggpairs(prestige_trim, aes(colour = type, alpha = 0.5), progress = FALSE)
```

High correlation, high significance with Education.
High correlation, high significance with Income
low negative correlation with women, not marked as significant
there seems to be a visible correlation between type and prestige: bc almost exclusively bottom half, prof top, wc sort of middle but skews bottom

```{r}
mod1a <- lm(prestige ~ education,
            data = prestige_trim)

summary(mod1a)
autoplot(mod1a)
```

```{r}
mod1b <- lm(prestige ~ type,
            data = prestige_trim)

summary(mod1b)
autoplot(mod1b)
```

prestige ~ education (mod1a):
r² = 0.7508
std error = 8.578 
predictor is statistically significant: p < 0.001 

paired plots mostly ok, bit of a u-shape in RvF

prestige ~ type (mod1b):
r² = 0.6976
std error = 9.499 
predictor is statistically significant across all levels: - p < 0.001
                                                            p < 0.001
                                                            p < 0.01

paired plots mostly ok, Q-Q maybe a little erratic towards the top?

education is a larger contributor, so makes more sense as the place to start.

model:

prestige = b0 + b1 * education
prestige = -10.8409 + 5.3884 * education

jobs start at negative prestige, but score increases by 5.38 per year of education

```{r}
prestige_remaining_resid <- prestige_trim %>% 
    add_residuals(model = mod1a) %>% 
    select(-c(prestige, education))
```

```{r message=FALSE}
prestige_remaining_resid %>% 
    ggpairs(aes(colour = type, alpha = 0.5))
```

significant strong positive correlation between residuals and income
significant strong negative correlation between women and income

```{r}
mod2a <- lm(prestige ~ education + income,
            data = prestige_trim)

summary(mod2a)
autoplot(mod2a)
```

```{r}
mod2b <- lm(prestige ~ education + type,
            data = prestige_trim)

summary(mod2b)
autoplot(mod2b)
```

again type is competitive, but not as much as its rival. 

prestige ~ education + income (mod2a):
r² = 0.814
std error = 7.45  
predictors are statistically significant: p < 0.001 for both

paired plots mostly ok, bit of a negative trend in the second half of RvF

prestige ~ education + type (mod1b):
r² = 0.7975
std error = 7.814  
predictors have issues in significance: education remains p < 0.001, but the only significant level of job type is now wc

paired plots mostly ok, little hump in RvF

income by far a better addition to the model.

model formula:


prestige = b0 + (b1 * education) + (b2 * income)
prestige = -7.621 + (4.292 * education) + (0.001 * income)

base prestige of -7.6, increases by 4.292 per year of education and by 0.001 per dollar of income

```{r}
prestige_remaining_resid <- prestige_trim %>% 
    add_residuals(mod2a) %>% 
    select(-c(education, income, prestige))
```

```{r}
ggpairs(prestige_remaining_resid, aes(colour = type, alpha = 0.5))
```

small negative correlation between women and the remaining residuals

```{r}
mod3a <- lm(prestige ~ education + income + women,
            data = prestige_trim)

summary(mod3a)
autoplot(mod3a)
```

```{r}
mod3b <- lm(prestige ~ education + income + type,
            data = prestige_trim)

summary(mod3b)
autoplot(mod3b)
```

Type might actually win!

prestige ~ education + income + women (mod3a):
r² = 0.8144 (effectively no change)
std error = 7.482   
women is not a statistically significant addition to the predictors (p = 0.6679)

paired plots mostly ok, bit of a wiggle in RvF

prestige ~ education + income + type (mod3b):
r² = 0.8349 (+0.02)
std error = 7.095   
predictors have issues in significance: now of the individual levels are significant, in fact bc (intercept) is p > 0.9!

paired plots mostly ok, bit of a curve in RvF

```{r}
summary(mod3b)
anova(mod2a, mod3b)
```
Analysis of Variance telling me that type is in fact significant (p < 0.01), so can justify being included in the model

New model formula:

bc is the reference class, which is now the intercept (b0)

prestige = -0.6229292 + (3.6731661 * education) + (0.0010132 * income) + (6.0389707 * (type == prof)) + (-2.7372307 * (type == wc))

```{r}
prestige_resid <- prestige_trim %>% 
    add_residuals(mod3b) %>% 
    select(-prestige)
```

```{r message=FALSE}
coplot("education":"income", data = prestige_resid)
```

