---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggfortify)
library(GGally)
library(fastDummies)
```

```{r}
house <- read_csv("data/housing_prices.csv")
```



```{r}
house <- house %>%
    mutate(rooms_per_house = total_rooms / households,
           people_per_house = population / households,
           bedrooms_per_house = total_bedrooms / households,
           bedrooms_per_person = total_bedrooms / population,
           rooms_per_person = total_rooms / population,
           # non_bedrooms = total_rooms - total_bedrooms,
           prop_bedrooms = total_bedrooms / total_rooms
    )
```

```{r}
house %>% skimr::skim() %>% View()
```

Observation: lots of the "count" data - non-zero, integer stuff - is highly positively skewed - we should consider log transformations if we're doing linear regression

```{r}
house_log <- house %>%
    select(-c(latitude, longitude, ocean_proximity)) %>%
    mutate(across(everything(), log)) %>%
    rename_with(~ paste0("log_", .x)) %>%
    bind_cols(house)
```

We now have a dataset with a richer feature set.
Lets start looking for relationships between variables, especially between our response variable and the potential predictors.

We can use ggpairs for this.

```{r}
house_log <- house_log %>%
    mutate(ocean_prox_sarah = if_else(
        ocean_proximity %in% c("ISLAND","NEAR BAY","NEAR OCEAN"), "NEAR WATER",
        ocean_proximity),
        ocean_prox_jamie = if_else(
            ocean_proximity %in% c("<1H OCEAN","NEAR BAY","NEAR OCEAN"), "NEAR WATER",
            ocean_proximity))
```

ggpiars might give more ideas for variable engineering!
it's a loop, not a straight road.

```{r message=FALSE}
house_log %>%
    select(median_house_value, housing_median_age, median_income, 
           ocean_prox_sarah, ocean_prox_jamie) %>%
    ggpairs(aes(fill = ocean_prox_sarah, alpha = 0.5))

# ggsave("ggpairs1.png")
```
we can use ggpairs to understand the relationships better and start making candidates for what might be the best predictors

early candidates: income, ocean_prox group, ..., ?age

```{r message=FALSE, warning=FALSE}
house_log %>% 
    select(median_house_value,
           log_bedrooms_per_house,
           log_households,
           log_total_bedrooms,
           log_prop_bedrooms,
           log_bedrooms_per_person,
           log_population,
           ocean_prox_sarah) %>% 
    ggpairs(aes(fill = ocean_prox_sarah,
                colour = ocean_prox_sarah,
                alpha = 0.5))
```

Let's start by finding the best single-predictor model for log(median_house_price)
Candidates: log(_)median_income, ocean_prox(_group)

```{r}
model1a <- lm(log(1 + median_house_value) ~ log(1 + median_income),
data = house_log)

summary(model1a)
```

```{r}
mod1b <- lm(median_house_value ~ ocean_proximity,
            data = house_log)

summary(mod1b)
```


