---
title: "Facets"
output: html_notebook
---

## Facetting
We convert one plot into multiple related plots
- helps ease the clarity of data
- provides a convenient and quick way to group data in related categories

```{r}
library(tidyverse)
```

```{r}
chick_weight <- janitor::clean_names(ChickWeight)
```

```{r}
ggplot(chick_weight) +
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  facet_wrap(~ diet, ncol = 4)
```

```{r}
benefits <- read_csv("data/benefits.csv")
```
Use geom_histogram to create a histogram showing the number of unemployed by age. Then facet the graph based on biological sex.
Hint: geom_histogram is almost the same as geom_bar, but is better suited to plotting continuous data because of the binning.

```{r}
ggplot(benefits) +
  aes(x = age, fill = sex) +
  geom_histogram() +
  facet_wrap(~ sex)
```

## `facet_grid()`

Pretty similar to `facet_wrap()` but gives more control when plotting two or more variables
x*y plots, even if some plots are empty

```{r}
ggplot(benefits) +
  aes(x = age, fill = sex) +
  geom_histogram(col = "white") +
  facet_wrap(married ~ sex)
```

```{r}
ggplot(benefits) +
  aes(x = age, fill = sex) +
  geom_histogram(col = "white") +
  facet_grid(married ~ sex)
```

Use the same dataset, but base the facet_grid() on marital status (married) and if the individual has children (dkids).
Extension: can you improve the way that the facets are labled.

```{r}
ggplot(benefits) +
  aes(x = age, fill = sex) +
  geom_histogram(col = "white") +
  facet_grid(married ~ dkids, labeller = "label_both")
```

## scales and space

```{r}
ggplot(benefits) +
  aes(x = age, fill = sex) +
  geom_histogram(col = "white") +
    facet_wrap(~ sex, ncol = 2, scales = "free_y")
```

