---
title: "Intro to ggplot"
output: html_notebook
---

```{r}
library(tidyverse)

students <- read_csv("data/students.csv")
```

```{r}
ggplot(data = students) + 
      aes(x = superpower) + 
      geom_bar()
```

```{r}
ggplot(students) +
      aes(x = superpower) +
      geom_bar(fill = "light blue", colour = "red")
```
```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar()
```

```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar(position = "dodge")
```

```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar(position = "fill")
```

## Statistical transformations

"count" is the default.

```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar(stat = "count")
```

```{r}
count_data <- students %>% 
  group_by(superpower, school_year) %>% 
  summarise(count = n())

ggplot(count_data) +
  aes(x = superpower, y = count, fill = school_year) +
  geom_bar(stat = "identity")
```

## Labels

```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar() +
  labs(
    x = "\nSuperpower",
    y = "Count",
    title = "Preferred Superpower by School Year",
    subtitle = "Data from students around the world",
    fill = "School Year"
  )
```

```{r}
ggplot(students) +
  aes(x = superpower, fill = school_year) +
  geom_bar() +
  xlab("\nSuperpower") +
  ylab("Count") +
  ggtitle("Preferred Superpower by School Year",
          subtitle = "Data from students around the world\n") +
  labs(fill = "School Year")
```

```{r}
olympics_overall_medals <- read_csv("data/olympics_overall_medals.csv")
```

```{r}
top_10 <- olympics_overall_medals %>%
  filter(season == "Summer", medal == "Gold") %>%
  arrange(desc(count)) %>%
  top_n(10)

top_10
```

```{r}
ggplot(top_10) +
  aes(x = team, y = count) +
  geom_bar(stat = "identity", fill = "gold") +
    labs(
    x = "Team",
    y = "Medal Count",
    title = "Gold medals by Olympic Team (Top 10)\n",
    fill = "Team")
```
# Layers

```{r}
library(janitor)
```
```{r}
chick_weight <- clean_names(ChickWeight)
```

```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet))
```
```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet)) +
  geom_point(aes(x = time, y = weight, colour = diet))
```

```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet)) +
  geom_point(aes(x = time, y = weight, colour = diet),
             alpha = 0.3) +
  geom_smooth(aes(x = time, y = weight, colour = diet))
```
```{r}
ggplot(chick_weight) +
  aes(x = time, y = weight, colour = diet) +
  geom_line(aes(group = chick)) +
  geom_point(alpha = 0.3) +
  geom_smooth()
```

Task 1 - 10 mins
Go back to using the students dataset:
Use geom_point to make a scatter graph, with the height of students on the x-axis and their reaction time of the y axis.
Make all the points blue. For geom_bar, the colour of the bar is controlled by fill, but for geom_point the colour of the points are controlled by colour.
Make the colour of the points depend on the superpower the student wishes they had.
Write down what the graph tells you overall.

```{r}
ggplot(students) +
  aes(x = height_cm, y = reaction_time, colour = superpower) +
  geom_point(fill = "Blue", size = 5) +
  labs(title = "Student Height vs Reaction Time",
       y = "Reaction time",
       x = "Height",
       colour = "Superpower")
```

Different layers can use different datasets, specified using `data` argument in a `geom`

```{r}
pets <- read_csv("data/pets.csv")
```


```{r}
ggplot(pets) +
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal), size = 4) +
  geom_text(
    aes(label = name),
    nudge_x = 0.5,
    nudge_y = 0.15,
    data = subset(pets, name == "Fluffy")
  )
```

# Saving

`ggsave` saves the last image by default

```{r}
ggplot(pets) +
  aes(x = age, y = weight) +
  geom_point(aes(colour = sex, shape = animal), size = 4) +
  geom_text(
    aes(label = name),
    nudge_x = 0.5,
    nudge_y = 0.15,
    data = subset(pets, name == "Fluffy")
  )

ggsave("g1_sav.pdf")
ggsave("g1_sav.png")
```