---
title: "Week 2 Day 4"
output: html_notebook
---

```{r}
library(tidyverse)
library(testthat)
```

# Unit Testing

Function we're going to test:

```{r}
is_leap_year <- function(year){
  
  ## if year mod 100 = 0
  if (year %% 100 == 0){
    ## if year mod 400 = 0
    if (year %% 400 == 0){
      return(TRUE)
    } else{
      return(FALSE)
    }
  } 
    
  ## if year mod 4 = 0
  if (year %% 4 == 0){
    return(TRUE)
  } else {
    return(FALSE)
  }
  
}
```

Function returns true if year mod 400 or year mod 4 = 0, otherwise returns false.

## `testthat()`

test_that("description",{
  test_1
  test_2
  ...
})

```{r}
test_that(desc = "Leap years return as TRUE", {
  expect_equal(is_leap_year(2020), TRUE)
  expect_equal(is_leap_year(2000), TRUE)
})
```
```{r}
# second test : one expectation
test_that("Character input returns an error", {
  expect_error(is_leap_year("2000"))
})
```

Tests that fail

```{r}
test_that("Non-numeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE))
})
```

TRUE reads as the number 1 if used in calculations, so we need to edit the function

```{r}
is_leap_year <- function(year){
  
  if(!is.numeric(year)){
    stop("Non-numeric input")
  }
  
  ## if year mod 100 = 0
  if (year %% 100 == 0){
    ## if year mod 400 = 0
    if (year %% 400 == 0){
      return(TRUE)
    } else{
      return(FALSE)
    }
  } 
    
  ## if year mod 4 = 0
  if (year %% 4 == 0){
    return(TRUE)
  } else {
    return(FALSE)
  }
  
}
```

```{r}
test_that("Non-numeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE))
})
```

## When should you use unit testing?

-- If you find yourself writing lots of informal tests
-- for extra documentation - tests give information about how your function should run
-- If you're writing a particularly challenging bit of code

Not essential when using pre-tested functions (like those in the tidyverse)

Test-Driven development

Red -> Green -> Refactor

Code initially doesn't pass tests -> RED
Write nasty code to pass tests -> GREEN
refactor code to be pretty but still pass tests -> REFACTOR

# Assertive Programming

Same analysis + same data = same results

We'll often want to carry out the same analysis on new data. New data might not be in quite the same format, leading to errors.

Assertive code tries to remedy this.

## Example -- Delhi Weather

```{r}
weather_one <- read_csv("data/delhi_weather_one.csv")

glimpse(weather_one)
```

```{r}
weather_report <- function(weather){
  # Our average temperature, humidity and wind speed 
  weather_means <- 
    weather %>% 
    summarise(
      mean_temp = mean(temp),
      mean_humidity = mean(humidity),
      mean_windspeed = mean(wind_speed)
    )
  
  # Show hot and humid days
  hot_and_humid <- 
    weather %>% 
    filter(temp > 20) %>% 
    filter(humidity > 55)
  
  return(list(weather_means, hot_and_humid))
}

weather_report(weather_one)
```

Now let's try our hopefully reproducible function on some new data

```{r}
weather_two <- read_csv("data/delhi_weather_two.csv")
```
```{r}
weather_report(weather_two)
```

## `assertr`

```{r}
library(assertr)
```

some values seem inappropriate

temperature much higher than recorded values

51C was the highest ever recorded
-34C was the lowest

we might want to limit our results table to include valid data only: temperatures between -35 and 55 deg C, for example

### `verify()`

```{r}
weather_report <- function(weather){
  
  # put in our checks first
  weather %>% 
    verify(humidity >= 0 & humidity <= 100) %>% 
    verify(temp >= -35 & temp <= 55) %>% 
    verify(wind_speed >= 0 & wind_speed <= 110)
  
  # Our average temperature, humidity and wind speed 
  weather_means <- 
    weather %>% 
    summarise(
      mean_temp = mean(temp),
      mean_humidity = mean(humidity),
      mean_windspeed = mean(wind_speed)
    )
  
  # Show hot and humid days
  hot_and_humid <- 
    weather %>% 
    filter(temp > 20) %>% 
    filter(humidity > 55)
  
  return(list(weather_means, hot_and_humid))
}

```

When verify throws an error, it pushes a data frame containing rows that failed verification

```{r}
weather_report(weather_two)
```

### `stopifnot()`

```{r}
weather_report <- function(weather){
  
  # put in our checks first
  stopifnot(
    nrow(weather) > 0,
    ncol(weather) == 4,
    c("temp", "humidity", "wind_speed" %in% names(weather_two))
  )
  
  # Our average temperature, humidity and wind speed 
  weather_means <- 
    weather %>% 
    summarise(
      mean_temp = mean(temp),
      mean_humidity = mean(humidity),
      mean_windspeed = mean(wind_speed)
    )
  
  # Show hot and humid days
  hot_and_humid <- 
    weather %>% 
    filter(temp > 20) %>% 
    filter(humidity > 55)
  
  return(list(weather_means, hot_and_humid))
}
```

```{r}
weather_report(weather_two)
```


