---
title: "Week 2 Day 4 - Reproducibility"
output: html_notebook
---

```{r libraries}
library(tidyverse)
```

# Rules for reproducibility

- Don't change the raw data! Treat it as read-only
- R-specific : work in Projects!
- follow a logical file naming format: lead with e.g 1_ if you need things to be done in order
- Flexible code and automated output

# Flexible code

```{r}
# Select species column

iris %>% 
  select(5)

# This is more flexible to changes in the data, and also more readable

iris %>% 
  select(Species)
```

# Functions and modular code

For functions with repeated use across multiple files, rather than creating the function in each file (at the top), we can create it once and load it into each file using `source()`


```{r}
source("subset.R")
```

```{r}
subset_function(1:100)
```

```{r}
source("01_multiply_function.R")
source("02_add_function.R")
```

```{r}
multiply_by_n(2, 4) %>% 
  add_two()
```

# Documentation

Include
- an outline of the purpose of the project
- instructions if people want to do the project / perform the analysis themselves
- discussion of assumptions: it's good to have a record of any assumptions you made
- dates / version history: I did this with these packages at this time.