---
title: "Probability"
output: html_notebook
---

```{r}
library(tidyverse)
library(gtools)
source("prob.R")
```

```{r}
s_two_dice <- rolldie(2)
```

```{r}
s_two_dice %>% 
  summarise(size_of_s = n())
```
Extract how many of our combinations lead to a total of 7

```{r}
s_sum_equals_seven <- s_two_dice %>% 
  filter(X1 + X2 == 7)
```

```{r}
s_two_dice <- s_two_dice %>% 
  mutate(total = X1 + X2)
```

We can use the new random variable `total` to define various event spaces

```{r}
s_sum_equals_seven <- s_two_dice %>% 
  filter(total == 7)
```


```{r}
s_sum_greater_than_10 <- s_two_dice %>% 
  filter(total > 10)
```

```{r}
s_two_dice <- s_two_dice %>% 
  mutate(
    first_die_even = (X1 %% 2 == 0),
    die2_minus_die1 = (X2 - X1),
    differ_by_two = case_when(
      X1 - X2 == 2 | X2 - X1 == 2 ~ TRUE,
      TRUE ~ FALSE)
  )
```

```{r}
s_two_dice %>% 
  filter(first_die_even == TRUE)
```

# Counting Rules

A way to figure out the number of outcomes in a probability problem.

## Single step experiments

Throwing two dice at the same time is a single step experiment.

All outcomes are equally likely - no control over how the die rolls.

prob(event) = num(event outcomes) / num(all)

```{r}
# get all combinations of 2 die rolls
s <- rolldie(2)

total_seven <- s %>% 
  filter(X1 + X2 == 7)
```

Now, using the above equation, we can work out the prob of getting a set of die rolls
that total 7

```{r}
(prob_total_seven <- nrow(total_seven) / nrow(s))
```
## Multiple step experiments

Let's say we throw one die, see the results, then flip a coin

If an experiment can be visualised as a sequence of N steps, with n1 possible outcomes
on the first step (e.g. 6), n2 possible outcomes on the second step (e.g. 2 : heads or tails)

We simply multiply the number of outcomes at each step to get the total number of outcomes.

n_total = n1 * n2 * ...

This means the above example has 12 total outcomes.

### Why do we need to know this?

The number of total outcomes can change the probability calculation.

e.g. What is the probability of getting heads on a coin toss and a 6 on a die throw?

- prob of heads = 0.5 or 1/2
- prob of getting a 6 in a die roll = 1/6
- prob of both events happening together = 1/2 * 1/6 = 1/12

```{r}

# set up one die roll
s_die <-  rolldie(1)

# set up one coin toss
s_coin <- tosscoin(1)

# merge them together
(s_total <- merge(s_coin, s_die))

# Q: Heads and a 6
(head_six <- s_total %>% 
    filter(toss1 == "H",
           X1 == 6))

# calculate the probability
(prob_head_six <- nrow(head_six) / nrow(s_total))
```

# Permutations

A permutation is an arrangement of objects in which **order matters**

e.g. locker code
Your code is 5243 - if you enter 3452 it wouldn't open because order matters
5243 is a different **permutation** than 3452

suppose you've remembered that your locker code involves the numbers
2, 3 , 4 and 5 but you can't remember the specific order

Q: how many permutations exist of these 4 numbers so you can figure out how long you'll have to spend trying to break into your locker.

Counting rule:
In plain language, it's "The number of permutations of N things taken r at a time"

N! / (N - r)!

! is a `factorial`
e.g. 5! = 5 * 4 * 3 * 2 * 1

e.g. Company hiring

Imagine we are filling 3 senior company positions (CEO, CTO, COO) from a board
of 10 employees

These 10 employees are pretty much equal in competence, and so for the sake of time we want to choose 3 of them **randomly** to fill the positions

Q: How many different permutations of possible employees are there that could end up making the 3 senior positions?

Note: counting rules exist so we don't have to list all the different possible outcomes.

```{r}
# create a vector of choices containing ten employees, n = 10

employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra", "Hua", "Inigo", "Hugo")

# Choose three at a time, r = 3
# Don't replace an employee once chosen: repeats.allowed = FALSE
# permutations = order matters, combinations = order doesn't matter

seniors <- as_tibble(gtools::permutations(n = 10, 
                                          r = 3, 
                                          v = employees, 
                                          repeats.allowed = FALSE)) %>% 
  rename(CEO = V1,
         CTO = V2,
         COO = V3)

# how many permutations do we have?

nrow(seniors)
```

This leads us nicely to asking an interesting analysis question from our data:

** what is the probability of Alice being selected as CEO and Bob being selected
as CTO, if the selection process is random? **

```{r}
# first we'll get the event space
alice_CEO_bob_CTO <- seniors %>% 
  filter(CEO == "Alice",
         CTO == "Bob")

# how many permutations do we have?

nrow(alice_CEO_bob_CTO)

# calculate the prob
nrow(alice_CEO_bob_CTO) / nrow(seniors)
```

This tells us there is a prob of 0.01 of Alice being CEO and Bob being CTO together

# Combinations

Combinations differ from permutations in that **the order of objects does not matter**

e.g lottery numbers
you select numbers, they're drawn, but the order of your numbers doesn't matter

A quick guide:
- permutations: you draw objects from a collection and place them into a group
**with** internal structure (e.g. hierarchy: CEO, CTO, COO)
- combinations: you draw objects from a collection and place them into a group 
**without** an internal struction, i.e. no labelled positions, roles or precedence
in the group

## Lottery example

Q1: Given a lottery with 50 numbers drawn in lots of 6, how many possible combinations
are there?

Q2: If you buy 10 tickets, each with a different set of numbers, what is your prob
of winning?

```{r}
# use choose() function, which just computes the number of choices from 
# 50 numers, drawn in ltos of 6

num_draws <- choose(50, 6)

# then we can calculate the prob as follows
prob_win_per_ticket <- 1 / num_draws
prob_win_per_ticket

# what if we have 10 tickets?
prob_win_per_ticket * 10
```

Task - 5 mins

Use the combinations() function in gtools for this task (don't try to do it manually). I have five flavours of ice cream to choose from:

flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio") 
and a bowl that can hold three scoops. How many different variations in bowls of ice cream can I have assuming:

Each scoop must be a different flavour
I can repeat flavours if I wish?

```{r}
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio") 

ice_cream <- as_tibble(combinations(n = 5, 
                                    r = 3, 
                                    v = flavours, 
                                    repeats.allowed = TRUE))

nrow(ice_cream)
```

# Combining Events

- Understand the concept of the complement of an event
- Be able to form 'AND' and 'OR' combinations of events
- Understand what it means for events to be **mutually exclusive**

## Complement

What happens if we want to know the likelihood of all observations **other** than
the one we're considering? This is the complement of an event

- Flip a coin, event is heads, complement is tails
- Pick a card from a deck and get spades suit, the complement is
{hearts, clubs, diamonds}

### Calculating the complement of an event

prob(roll = 4) = 1/6
prob(not rolling a 4) = 5/6

### Examples

Toss 4 coins

```{r}
s_four_coins <- tosscoin(4) %>% 
  mutate(row = row_number())
```

What is the probability of the event that at least 2 heads are thrown?

```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_heads = rowSums(. == "H"))

# now we can filter for where we have 2 heads thrown and calculate the prob

s_at_least_two_heads <- s_four_coins %>% 
  filter(num_heads >= 2)

# calculate the prob

prob_two_heads <- nrow(s_at_least_two_heads) / nrow(s_four_coins)
```
What about the **complement**? In this case, fewer than 2 heads.

```{r}
s_less_than_two_heads <- s_four_coins %>% 
  filter(num_heads <)

# prob

prob_less_two_heads <- nrow(s_less_than_two_heads) / nrow(s_four_coins)
```

```{r}
prob_two_heads + prob_less_two_heads
```

As datasets grow, and calculations become more complex, the complement rule
can save you time (1 - calculated_probability)

# AND combinations

"Compound event" -- just an event composed of a combination of other events.

e.g. roll a single die
  - Event A: one of {1, 2, 3, 4} is rolled
  - Event B: one of {3, 4, 5, 6} is rolled
  
What is the prob that both A **and** B occur?

The conditions that which both happen is called the "intersection". 
In maths the symbol is ∩

For equally likely outcomes, the prob of A **and** B is just defined the usual way:
the num of outcomes in the intersection, divided by the size of the sample space

prob(A AND B) = num(A AND B) / num(S) = 2/6 = 1/3

Task - 5 mins

Given our four coin tossing example from above, what is the probability of tossing at least one head and at least one tail?

To do this, you'll need to add a num_tails column to s_four_coins
Then think about filtering to satisfy the two conditions

```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_tails = rowSums(. == "T"))
```

```{r}
s_one_of_each <- s_four_coins %>% 
  filter(num_tails >= 1 & num_heads >= 1)

prob_one_of_each = nrow(s_one_of_each) / nrow(s_four_coins)
```

## Mutually exclusive events

What about the case that two (or more) events have no outcomes in common?
In this case, the prob of the intersection is zero, **it is not possible for the
events to occur together**

e.g. 4 coins tossed: prob of at least 3 heads **and** at least 2 tails

```{r}
s_four_coins %>% 
  filter(num_heads >= 3 & num_tails >= 2) %>% 
  summarise(prob = n() / nrow(s_four_coins))
```

The events (3+ heads and 2+ tails) are **mutually exclusive** when we flip 4 coins

# OR combinations

e.g. roll a single die
  - event A: roll an even number {2, 4, 6}
  - event B: roll one of {4,5}
  
  What is the prob that A **or** B occurs?
  
The condition where this happens is called the `Union` and is represented by ∪ in maths

----

Let's do it with two dice. To


Experiment consists of rolling 2 die rolls and defining the two events.

```{r}
s_two_dice <- rolldie(2)

# filter down to cases where X1 is even or X2 is 4 or 5

s_two_dice %>% 
  filter(X1 %% 2 == 0 | (X2 == 4 | X2 == 5)) %>% 
  summarise(prob = n() / nrow(s_two_dice))
```

# Conditional probability

So far, we've looked at the prob of one or two events happening totally apart from each other e.g. two coin flips, dice rolls etc.

But more often we'd like to be able to calculate the prob of one event **given that**
another has occurred

e.g. what are the chances of selling item X **if** we've already sold item Y

We call this **conditional probability** - written prob(A | B)

e.g. "What is the prob that a 5 was rolled **given that** we know an odd number was rolled?"

In this case, the event space contains 3 outcomes: {1,3,5}

prob(5 | odd) = 1/3

The **conditioning event** (i.e. number is odd in this case) **filters down**
the space of our outcomes we have to consider.

## Dependent and independent events

We say that 2 events A and B are **independent** if
  prob(A | B) = prob(A)

i.e. the prob of event A is **not affected** by whether event B has happened or not

Example: colour blindness data

```{r}
cb <- read_csv("data/colour_blindness.csv")
```
Can view this simple categorical data set with a **contingency table**

```{r}
table <- cb %>% 
  janitor::tabyl(bio_sex, cblind)
```

```{r}
table2 <- table %>% 
  janitor::adorn_totals(where = c("row", "col"))
```

If we prefer to see this data in terms of relative proporations, we can use
adorn_percentages() function

```{r}
table3 <- table2 %>% 
  janitor::adorn_percentages(denominator = "all") %>% 
  janitor::adorn_title()
```

## Marginal probabilities

The numbers in the Total row and col correspond to what are called
**marginal probs** (because they appear at the table margins)

If we didn't have access to this contingency table, we could calculate the 
prob(female) using pipes:

```{r}
prob_female <- cb %>% 
  summarise(prob = sum(bio_sex == "female") / n()) %>% 
  pull()
```

Task - 5 mins

What are the marginal probabilities of:

being male
being colour blind
not being colour blind?
Get these probabilities from the contingency table, and also calculate them using pipes. Save each probability in a sensibly named variable, we'll need it later!

```{r}
prob_male <- cb %>% 
  summarise(prob = sum(bio_sex == "male") / n()) %>% 
  pull()

prob_cb <- cb %>% 
  summarise(prob = sum(cblind) / n()) %>% 
  pull()

prob_not_cb <- cb %>% 
  summarise(prob = sum(!cblind) / n()) %>% 
  pull()
```

## Marginal probabilities

The probs in the main body of the table are called **joint probs** and they correspond
to the AND combination of the corresponding levels

```{r}
prob_female_cb <- cb %>% 
  filter(bio_sex == "female",
         cblind == TRUE) %>% 
  summarise(prob = n() / nrow(cb)) %>% 
  pull()

prob_male_cb <- cb %>% 
  filter(bio_sex == "male",
         cblind == TRUE) %>% 
  summarise(prob = n() / nrow(cb)) %>% 
  pull()
```

## Dependent and independent events

Is the event 'being female' dependent or independent of the event 'being colourblind'?

```{r}
prob_female
prob_female * prob_cb
```

These values are unequal! This means that the two events (being female and being
colour blind) are **dependent**

```{r}
prob_male
prob_male * prob_cb
```

## Conditional probability

The easiest way to see the dependence of events if by calculating their
**conditional probabilities** 
in this case: What is the prob of a person being colour blind **given** that
we know they are female?

i.e. what proportion of female people are colour blind?
  prob(colour blind | female)
  = prob (colour blind AND female) / prob(female)
  = 0.0016 / 0.4
  = 0.004
  = 0.4%

```{r}
prob_cb_given_female <- cb %>% 
  filter(bio_sex == "female") %>% 
  summarise(prob = sum(cblind == TRUE) / n())

prob_cb_given_female
```

```{r}
prob_cb_given_male <- cb %>% 
  filter(bio_sex == "male") %>% 
  summarise(prob = sum(cblind == TRUE) / n())

prob_cb_given_male
```

The prob of an individual being collur blind **depends** upon their biological
sex. This is true for this set of 10,000 individuals - what about more generally?

If this 10,000 was a sample of a broader population, are prob(colourblind | female)
and prob(colourblind | male) sufficiently different in this sample to convince us
that this is also likely to be a real effect in the broader population?

Covered later in statistical testing.