---
title: "Sampling Distributions"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)

telco <- read_csv("data/telecomms_churn.csv") %>% 
  clean_names()
```

```{r}
skimr::skim(telco)
```


For the purposes of this lesson, let's pretend this is the entire population, which we hardly ever get to see in real life.


```{r}
summary_pop <- telco %>% 
  summarise(mean_monthly_charges = mean(monthly_charges),
           mean_tenure = mean(tenure),
           median_tenure = median(tenure),
           prop_churn = sum(churn == "Yes") / n())
```

```{r}
telco %>% 
  ggplot(aes(x = monthly_charges)) +
  geom_histogram(col = "white")
```

```{r}
telco %>% 
  ggplot(aes(x = tenure)) +
  geom_histogram(col = "white") +
  scale_x_continuous(breaks = seq(0, max(telco$tenure), 6))
```

```{r}
telco %>% 
  ggplot(aes(x = churn)) +
  geom_bar(col = "white")
```

Now we'll simulate the circumstances of having small samples of data from this population, so that we can see how large a role random variation plays.

```{r}
sample_200 <- telco %>% 
  rep_sample_n(200)

sample_200
```

```{r}
summary_sample_200 <- sample_200 %>% 
  ungroup() %>% 
  summarise(mean_monthly_charges = mean(monthly_charges),
           mean_tenure = mean(tenure),
           median_tenure = median(tenure),
           prop_churn = sum(churn == "Yes") / n())
```

"Sampling Error"

Difference between the point estimate (e.g. sample mean) and the population parameter (e.g. population mean)

```{r}
summary_sample_200 - summary_pop
```

```{r}
sample_200_n <- telco %>% 
  rep_sample_n(200, reps = 5000)
```

```{r}
summary_sample_200_n <- sample_200_n %>% 
  summarise(mean_monthly_charges = mean(monthly_charges),
           mean_tenure = mean(tenure),
           median_tenure = median(tenure),
           prop_churn = sum(churn == "Yes") / n())
```

```{r}
summary_sample_200_n %>% 
  ggplot(aes(x = mean_tenure)) +
  geom_histogram(col = "white")
```

```{r}
library(fastGraph)

std_errors <- summary_sample_200_n %>% 
  summarise(sd_mean_monthly_charges = sd(mean_monthly_charges))

shadeDist(xshade = c(60, 70),
          lower.tail = FALSE,
          ddist = "dnorm",
          parm1 = mean(summary_sample_200_n$mean_monthly_charges),
          parm2 = std_errors$se_mean_monthly_charges,
          xlab = "Mean Monthly Charges")

summary_sample_200_n %>% 
  filter(mean_monthly_charges >= 60,
         mean_monthly_charges <= 70) %>% 
  summarise(prob = n() / nrow(summary_sample_200_n))
```